### 2016-03-02 paper/ms.tex
Updated Table 1 with lots more measurements

%\documentclass[12pt,preprint]{aastex}
\documentclass{emulateapj}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{url}
\usepackage{bm}
\bibliographystyle{aasjournal}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\renewcommand{\vec}[1]{\bm{#1}}
\DeclareMathOperator\erf{erf}
\begin{document}

\title{The Inner Mass-Ratio Distribution of Intermediate-Mass Stars}

\author{Kevin Gullikson \altaffilmark{1}}
\author{Adam Kraus \altaffilmark{1}}
\author{Sarah Dodson-Robinson \altaffilmark{2}}

\altaffiltext{1}{University of Texas, Astronomy Department. 2515 Speedway, Stop C1400. Austin, TX 78712}
\altaffiltext{2}{University of Delaware, Department of Physics and Astronomy, 217 Sharp Lab, Newark, DE 19716}

\begin{abstract}
Binary stars and higher-order multiple systems are a ubiquitous outcome of star formation, especially as the system mass increases. The distribution of mass ratios is a unique probe into the conditions of the collapsing cloud core and circumstellar disk(s) of the binary fragments. Inside $a \sim 1000$ AU the disks from the two forming stars can interact, and additionally companions can form directly through disk fragmentation. We should therefore expect the mass-ratio distribution of close companions to differ from that of wide companions. This prediction is difficult to test with intermediate-mass to massive primary stars because close companions are so much dimmer. We present the results of a survey searching for companions to A- and B-type stars using the direct spectral detection technique, which is sensitive to late-type companions within $\sim 1''$ of the primary and which has no inner working angle. We estimate the temperatures and masses of the 57 companions , 25 of which are new detections. We find that the mass-ratio distribution for our sample has a robust maximum near $q \sim 0.3$. Our mass-ratio distribution has a very different form than in previous work, where it is usually well-described by a power law, and indicates that close companions to intermediate-mass stars experience significantly different accretion histories or formation mechanisms than wide companions.
\end{abstract}
\keywords{binaries: spectroscopic, stars:early-type, stars:formation, stars:statistics}

\maketitle

\section{Introduction}

\label{sec:intro}

Stellar multiplicity is an inevitable and common outcome of star formation, with roughly half of all solar-type field stars in binary or multiple systems \citep{Raghavan2010} and an even higher fraction as the stellar mass increases \citep{Zinnecker2007}. Young stellar associations and clusters tend to have even higher multiplicity \citep{Duchene2013}, indicating that stars often form in multiple systems that are subsequently destroyed by dynamical interactions as the cluster dissociates.

The overall multiplicity rate and the distributions of mass ratio, period, and eccentricity of a binary star population place important constraints on the mode of binary star formation. While the period and eccentricity are altered by dynamical processing in the birth cluster, the present-day mass ratio of a binary system is a direct result of its formation \citep{Parker2013}. Most binary stars are thought to form via core fragmentation \citep{Boss1979, Boss1986, Bate1995}, in which a collapsing core fragments into two or more individual protostars. The number and initial masses of the fragments are set by the total core mass, as well as its rotation, turbulence, and its temperature and density structure. If the fragments are well separated ($a \gtrsim 1000$ AU), they will evolve independently of each other, accreting mass from the core material onto their own protostellar disks and then onto the protostars themselves. However close fragments ($a \sim 100$ AU) will interact with each other; the protostellar disk may be truncated, destabilized, or form into a circumbinary disk if the separation is small enough \citep{Bate1997}. In addition, an unstable disk can fragment to form low-mass companions \citep{Kratter2006, Stamatellos2011}. The mass ratios of close companions formed via either mechanism should be affected by preferential accretion. Most work has suggested that the disk material will preferentially accrete onto the lower mass companion \citep{Bate1997, BBB2002}; however, recent work has indicated that magnetic disk braking may result in preferential accretion onto the more massive component \citep{Zhao2013} instead. In either case, we would expect to find a mass ratio distribution for companions inside a few $100$ AU that differs from that of companions on wider orbits.

The mass ratio, period, and eccentricity distributions are fairly well-known for solar type stars \citep{Duquennoy1991, Raghavan2010} and cooler stars \citep{Fischer1992, Delfosse2004}. Interestingly, the mass ratio distribution appears to be invariant to separation for these stars \citep{Meyer2013}, contrary to the theoretical expectations. All of the distributions are much less certain for more massive stars. The reason for this is two-fold: first, more massive stars tend to more rare and farther away, meaning many of the companions are angularly close to the very bright primaries and difficult to detect with imaging techniques. Second, the primary stars tend to be rapid rotators, which limits radial velocity precision to $\sim 1 \mathrm{km\ s}^{-1}$. Additionally, radial velocity monitoring can only measure a mass ratio if spectral lines from both components are visible; this typically suffers from the same flux ratio difficulty as imaging techniques.

Nonetheless, \citet{DeRosa2014} performed an adaptive optics imaging survey of nearby A-type stars, and found that the mass ratio distribution is well-described by a power law with large slope, indicating a very strong preference for low-mass companions. They also found initial evidence that the mass ratio distribution for companions inside $125$ AU has a much shallower power law slope than that of wide companions, and is consistent with flat. Their close companion subsample contained only 18 binary systems, and the result is complicated by the inherent difficulty of detecting close companions with low mass ratios in an imaging survey.

Radial velocity monitoring surveys can detect much closer companions than imaging surveys, but are typically only complete to low-mass companions if the primary is a slow rotator. Chemically peculiar Am stars are typically associated with binary companions, and are slow rotators due to tidal braking; they thus form a highly biased sample of intermediate-mass stars. Nonetheless, it is interesting to note that they have a mass-ratio distribution which peaks near $q \sim 0.5$ \citep{Vuissoz2004}, an entirely different form than the distribution found around chemically normal stars at wide separations.

In this paper, we describe a spectroscopic survey of nearby chemically normal, main sequence intermediate-mass stars ($M \approx 1.5 - 15 M_{\odot}$). We search for companions using the direct spectral detection technique \citep{Gullikson2016}, which has a separation-invariant detection rate for all separations inside $\sim 1 ''$. We describe the stellar sample and data used for the survey in Section \ref{sec:obs}, as well as the data reduction steps in the same section. Next, we describe the direct spectral detection method and tabulate the companion detections in Section \ref{sec:companions}. We estimate the mass and age of the primary stars in Section \ref{sec:sp}, and discuss the survey completeness in Section \ref{sec:completeness}. Finally, we end with a derivation of the mass ratio distribution from our sample in Section \ref{sec:mrd} and discuss its implications for binary formation in Section \ref{sec:discussion}.

\section{Observations and Data Reduction}
\label{sec:obs}

The stellar sample for this survey is defined by the following criteria:

\begin{itemize}
\item $V < 6$ mag
\item $v\sin{i} > 80 \mathrm{km\ s}^{-1}$
\item Spectral Type A or B with the following additional constraints
\begin{itemize}
  \item Main Sequence
  \item No spectral peculiarities except for `n', which denotes broad lines.
\end{itemize}
\end{itemize}

The magnitude limit ensures that a sufficiently high signal-to-noise ratio can be achieved in a short period of time. It does introduce a Malmquist bias in the derived mass ratio, which we discuss and correct for in Section \ref{sec:mrd}. Likewise, the $v\sin{i}$ limit makes accounting for the primary star spectrum in the companion search trivial; since most A- or B-type stars are rapid rotators, the cutoff removes less than half of the stars from the potential sample. We exclude pre-main sequence stars because both the primary and the companion mass would depend very strongly on young and uncertain ($\lesssim 1$ Myr) evolutionary models. Finally, we exclude post-main sequence stars from our sample because the binary flux ratio would be even less favorable to companion detection in an evolved star. Most of the spectral peculiarities denote narrow lines, which are already removed from the sample by the $v\sin{i}$ cut. The sample is given in Table \ref{tab:sample}. The spectral type, coordinates, V magnitude, and parallax are all adopted from the Simbad Database \citep{Simbad}.

\subsection{Spectroscopic Data}
\label{subsec:specdata}
We use several high spectral resolution, cross-dispersed \'echelle spectrographs for this survey. We use the CHIRON spectrograph \citep{CHIRON} on the 1.5m telescope at Cerro Tololo Inter-American Observatory for most southern targets. This spectrograph is an $R\equiv \lambda / \Delta \lambda = 80000$ spectrograph with wavelength coverage from 450 - 850 nm, and is fed by a $2.7''$ optical fiber. The data are automatically reduced with a standard CHIRON data reduction pipeline, but the pipeline leaves residuals of strong lines in adjacent orders. We therefore bias-correct, flat-field and extract the spectra with the optimum extraction technique \citep{Horne1986} using standard IRAF\footnote{IRAF is distributed by the National Optical Astronomy Observatories, which are operated by the Association of Universities for Research in Astronomy, Inc., under cooperative agreement with the National Science Foundation.} tasks, and use the wavelength calibration from the pipeline reduced spectra.

For the northern targets, we use a combination of the High Resolution Spectrograph \citep[HRS,][]{HRS} on the Hobby Eberly Telescope, and the Tull coud\'e \citep[TS23,][]{TS23} and IGRINS \citep{IGRINS} spectrographs, both on the 2.7m Harlan J. Smith Telescope. All three northern instruments are at McDonald Observatory. For the HRS, we use the $R = 60000$ setting with a $2''$ fiber, and with wavelength coverage from 410-780 nm. We bias-correct, flat-field, and extract the spectra using an IRAF pipeline very similar to the one we use for the CHIRON data. The HRS spectra are wavelength-calibrated using a Th-Ar lamp observed immediately before or after the science observations.

For the TS23 spectrograph, we use a $1.2''$ slit in combination with the E2 \'echelle grating (53 grooves/mm, blaze angle $65^{\circ}$), yielding a resolving power of $R=60000$ and a wavelength coverage from 375-1020 nm. We reduce the data using an IRAF pipeline very similar to the ones we use for CHIRON and HRS, and wavelength calibrate using a Th-Ar lamp observed immediately before the science observations.

\begin{figure*}
\includegraphics[width=\textwidth]{HIP_109139.pdf}
\caption{\emph{Left}: Cross-correlation function between the observed spectra of HIP 109139 and a $5700\ K$ Phoenix model spectra. The detection at two dates shows significant velocity variation, indicating orbital motion with a short period. \emph{Right}: Peak CCF height as a function of Phoenix model spectra template temperature. The maxima of the curves indicate the temperature of the companion.}
\label{fig:ccf}
\end{figure*}

IGRINS only has one setting with $R = 40000$. It has complete wavelength coverage from $1475-2480$ nm, except in the telluric water band from $1810 - 1930$ nm. Each star is observed in an ABBA nodding mode, and reduced using the standard IGRINS pipeline \citep{IGRINS_plp_v2}. The standard pipeline uses atmospheric OH emission lines as well as a Th-Ar calibration frame to calibrate the wavelengths; we further refine the wavelength solution using telluric absorption lines in the science spectrum.

After reducing the data, we fit and remove the telluric spectrum using the TelFit code \citep{Gullikson2014}. We fit each \'echelle order affected by telluric absorption independently from each other to get the best removal. The telluric correction is critical for IGRINS spectra, where every order is dominated by telluric absorption lines. For the optical spectra, it is less critical but allows us to use some of the redder orders than we otherwise would be able to. For unsaturated lines, the best-fit telluric model reproduces the data to within $\sim 1-5\%$ of the continuum level.

We give the spectroscopic observation log in Table \ref{tab:observations}. We calculate the signal-to-noise ratio (the ``snr'' column) for the optical instruments (CHIRON, TS23, and HRS) as the median of the extracted flux divided by it's uncertainty for each pixel from the \'echelle order nearest $675$ nm. For the IGRINS instrument, we calculate the signal-to-noise ratio from the order nearest $2200$ nm.

TODO: MAKE THE SPECTRA AVAILABLE SOMEHOW...

\subsection{Imaging Data}
As part of the follow-up effort, we used the NIRI instrument behind the Altair adaptive optics system at the Gemini North Telescope. For each star listed in Table \ref{tab:imaging_obs}, we obtained 25 images in 5 dithering positions. We used the K-continuum band centered on $2.2718\ \mu m$ and a variety of exposure times and dates (listed in Table \ref{tab:imaging_obs}). Because the targets are all extremely bright, we used the high read noise and high flux detector settings to allow for very short co-add exposure times. We reduced the data using the Gemini set of IRAF tasks, which include steps for nonlinearity correction, flat-fielding, sky subtraction, and co-addition of the dither frames.

We measure the flux and position of both stars by fitting a 2D Moffat function \citep{Moffat1969} to both stars simultaneously, constraining the shape parameters for both functions to be the same. The ratio of the amplitudes gives the magnitude difference, and the pixel locations along with the detector pixel scale gives the separation and position angle between the stars. We note that the goal of these images was confirmation and we did not observe any reference targets to make a distortion map and correct the image rotation. The uncertainty in position angle and to a lesser degree separation quoted in Table \ref{tab:imaging_obs} is likely underestimated.

\section{Companion Search}
\label{sec:companions}

We search for stellar companions to our sample stars using the direct spectral detection technique, described in detail in \citet{Gullikson2016}. In short, we unsharp-mask each spectrum using a gaussian filter with width proportional to the primary star $v\sin{i}$ to remove the broad lines from the primary star. We then cross-correlate each \'echelle order of each filtered spectrum against a large grid of Phoenix model spectra \citep{Husser2013_b} with the following parameters:

\begin{itemize}
\item $T_\mathrm{eff} = 3000-12000$ K, in steps of 100 K
\item {[}Fe/H{]} = -0.5, 0.0, +0.5
\item $v\sin{i} = 1, 5, 10, 20, 30 \ \mathrm{km\ s}^{-1}$
\end{itemize}

We combine the cross-correlation functions for all orders using both a simple average and the maximum-likelihood weighting scheme \citep{Zucker2003}. A companion detection is denoted by a strong peak in the combined cross-correlation function (CCF). While the maximum-likelihood scheme produces detections with much higher significance, it also magnifies spurious peaks and so has a larger false-positive rate. For this reason, we use the simple average CCFs in all further analysis.

The peak height in the CCF as a function of the stellar model acts in a similar way to the more typical $\chi^2$ map of parameter space. More concretely, as the stellar model template gets closer to the true companion spectrum, the CCF peak gets higher. We can therefore measure the companion temperature and, to a lesser degree its metallicity and $v\sin{i}$, in a single spectrum. Imperfect stellar models cause a bias between the true companion temperature and the temperature of the Phoenix model which produces the largest CCF peak. This bias is most pronounced at low temperatures, where the difficult-to-model molecular absorption becomes important. We correct for the bias by applying the calibrations developed in \citet{Gullikson2016}. These calibrations are only valid for companions with $3000 < T_\mathrm{eff} < 7000 K$; for detections at hotter temperatures we assume that the temperature which produces the maximum CCF peak is an \emph{unbiased} estimator of the true companion temperature.

We list the companion detections in Table \ref{tab:companions}, and report the estimate of the companion temperature, $v\sin{i}$, and metallicity derived from the model parameters which produce the largest CCF peak. The $v\sin{i}$ and metallicity values do not have uncertainties and should only be taken as a rough estimate of the true value. The mean and standard deviation of the companion metallicities is $-0.29 \pm 0.30$; the marginal bias towards low metallicities is most likely a measurement bias and does not reflect the true companion population \citep{Gullikson2016}.  We show the detection CCFs and a plot of peak CCF height as a function of model temperature for HIP 109139 in Figure \ref{fig:ccf}. Similar figures for all companions are available in the supplementary files.

We have follow-up spectroscopy for 17/27 of the new companions to confirm their existence. In most cases, there is a clear shift in the radial velocity of the companion, indicating that it is orbiting the target star and is not a foreground or background contaminant (See Figure \ref{fig:ccf}). Two of the new detections (companions to HIPs 38593 and 79404) were detected twice but not in a third attempt, most likely because the third spectrum had low signal-to-noise. The companion to HIP 93805, at $\sim 4000$ K, was detected twice with near-infrared IGRINS but not the optical CHIRON instrument that is less sensitive to cool companions than IGRINS. Two of the companions with only one detection were observed at least twice (HIPs 19949 and HIP 23362); both of the non-detections are from the IGRINS instrument, which is less sensitive to hot companions with rapid rotation speeds because there are far fewer spectral lines of the companion in the near-infrared than there are in the optical.

In addition we obtained Gemini/NIRI adaptive optics imaging data for 18 of the northern companions, and were able to resolve 7 of them. We show the separation, position angle, and magnitude difference measurements in Table \ref{tab:imaging_obs}, and display the images in Figure \ref{fig:images}. We also derive the projected separation in AU and the companion mass from the images. We calculate the separation from the measured angular separation and the Hipparchos parallax \citep{Hipparchos}. We calculate the companion mass and uncertainties from 1000 samples of the magnitude difference measurement and the primary star mass, temperature, age, and radius (see Section \ref{sec:sp}). For each sample, we use a grid of Kurucz stellar model spectra [CITE] and the pysynphot code [CITE] to determine the companion temperature needed to replicate the observed magnitude difference. We estimate the companion radius by interpolating solar metallicity Dartmouth isochrones [CITE] from the companion temperature and system age sample. We convert the best temperature to a companion mass using the same isochrone grid. The masses derived from the imaging data have very large uncertainties because the primary star property estimates that they depend on are very uncertain.

One star in the imaging sample, HIP 88116, has several nearby sources in the image. We quote the magnitude difference and separation of the brightest source in Table \ref{tab:imaging_obs}, but stress that \emph{none} of the visible sources is likely to be the companion we see in the spectroscopic data. The two epochs of spectroscopic data show a radial velocity shift of $\Delta v = 30.7\ \mathrm{km\ s}^{-1}$ over the course of roughly one year; this orbital motion is much too large to allow for any of the companions visible in the image (all with separations $ > 1''$ and projected separations $ > 300$ AU).

\begin{figure}
\includegraphics[width=\columnwidth]{Imaging_Data.pdf}
\caption{Detection images for all stars in which we detect a companion in the follow-up NIRI data. 
There are several nearby sources for HIP 88116, none of which are the source we detect in the 
spectroscopic data (see Section \ref{sec:companions}).}
\label{fig:images}
\end{figure}

\section{Primary Star Parameters}
\label{sec:sp}

In order to convert from companion temperature to mass ratio, we first need an estimate of the primary mass. In addition, since the primary stars in our survey have short main-sequence lifetimes, some companions may still be contracting onto the main sequence and so an age estimate for the system is necessary to convert from companion temperature to mass.

About half of our detections have robust mass and age estimates for the primary stars from Str\"omgren uvby$\beta$ photometry \citep{David2015}. For those that do not, we estimate the mass and age of the system from our spectra. We first cross-correlate the data against a grid of model spectra spanning

\begin{itemize}
\item $7000\ K < T_\mathrm{eff} < 30000\ K$ in steps of $1000\ K$
\item $3.0 < \log{g} < 4.5$ in steps of $0.5$ dex
\item $75 < v\sin{i} < 300\ \mathrm{km s}^{-1}$ in steps of $25\ \mathrm{km s}^{-1}$
\end{itemize}
Since the Phoenix model library does not extend to temperatures $T > 12000\ K$, we used Kurucz model spectra \citep{Castelli2003} for this grid.

For the optical data, we use the blue echelle orders ($\lambda < 5550 \AA$). We ignore the strong hydrogren Balmer lines in the spectrum because they span several echelle orders and make continuum normalization very difficult, potentially biasing the result. There are sufficient metal lines in the optical spectra that the resulting CCF always has a very strong peak at the radial velocity of the primary star. The near-infrared IGRINS spectra have very few strong metal lines; we use the subset from $1.51 - 1.73 \micron$ that is dominated by hydrogren Brackett lines for these spectra. We then estimate the temperature and surface gravity of the primary stars from the CCF with the largest peak. We adopt the following errors on the temperature and surface gravity:

\begin{align}
 \sigma_T &= \begin{cases}
      \hfill 300\ K \hfill & T < 10000\ K \\
      \hfill 1000\ K \hfill & T >= 10000\ K \\
     \end{cases} \\
 \sigma_{\log{g}} &= 0.25
\end{align}

Next, we use Padova stellar evolutionary tracks \citep{Bressan2012} and the isochrones code \citep{isochrones_code} to estimate the mass and age of the system from the measured temperature and $\log{g}$. As a consistency check, we also interpolate from a table of stellar properties as a function of spectral type \citep{Pecaut2013} to estimate the primary mass from the published spectral types. We show the comparison in Figure \ref{fig:prim_mass}. We estimate uncertainties in the spectral type mass by assuming a spectral type uncertainty of $\pm 0.5$ spectral types and propagating to mass. There is excellent agreement between the masses we measure and the spectral type masses.

\begin{figure}
\includegraphics[width=\columnwidth]{PrimaryMassEstimates_log.pdf}
\caption{Comparison of primary star masses derived from our cross-correlation analysis and Padova isochrones \citep{Bressan2012} with those expected from the published spectral type. There is excellent agreement between the two measures across the entire range of masses.}
\label{fig:prim_mass}.
\end{figure}

We show the temperature, surface gravity, mass, and age estimates for all primary stars in Table \ref{tab:sample}. The MCMC samples generated by the isochrones code are available at [GIVE ZENODO LINK OR SOMETHING].

\section{Survey Completeness}
\label{sec:completeness}

The detectability of a companion mostly depends on its temperature: cooler companions emit much less light and so are increasingly lost in the Poisson noise from the primary star spectrum. A companion with a high rotation rate is also more difficult to detect because the cross-correlation function gets most of its power from narrow spectral lines.

\subsection{Sensitivity Analysis}

To quantify the detection rate as a function of companion temperature and $v\sin{i}$, we performed a series of injection/recovery experiments. We started by creating synthetic binary star observations from each of our observed spectra. We made two distinct grids of companion stars: a low temperature grid spanning

\begin{itemize}
\item $3000\ K < T_\mathrm{eff} < 6500\ K$ in steps of $100\ K$
\item $0\ \mathrm{km\ s}^{-1} < v\sin{i} < 50\ \mathrm{km\ s}^{-1}$ in steps of $10\ \mathrm{km\ s}^{-1}$
\end{itemize}
and a high temperature grid spanning

\begin{itemize}
\item $7000\ K < T_\mathrm{eff} < 12000\ K$ in steps of $1000\ K$
\item $100\ \mathrm{km\ s}^{-1} < v\sin{i} < 250\ \mathrm{km\ s}^{-1}$ in steps of $50\ \mathrm{km\ s}^{-1}$
\end{itemize}
For each grid point, we added a solar metallicity Phoenix model spectrum to the observed data after scaling to replicate the expected flux between a main sequence companion of the model temperature and the known target star spectral type. If the target star had known companions within $3''$, we included the expected flux from the companion when computing the flux ratio. We repeated each grid point at different radial velocities spanning $-400\ \mathrm{km\ s}^{-1} < v < 400\ \mathrm{km\ s}^{-1}$ in $50\ \mathrm{km\ s}^{-1}$ steps to sample the noise properties of the spectra and estimate a probability of detection at each point.

Next, we cross correlated all of the synthetic observations against the Phoenix model template that was used to construct them. We counted the companion as detected if the highest point in the resulting CCF was found at the correct radial velocity. We then combined all of the radial velocity trials for each grid point to estimate a probability of detection at that grid point:

\begin{equation}
P(\mathrm{detection}) = \frac{N_\mathrm{detected}}{N_\mathrm{rv}}
\end{equation}
where $N_\mathrm{rv} = 17$ is the number of radial velocity trial points.

Finally, we interpolated between the grid points using a linear radial basis function interpolator (Figure \ref{fig:detrate_2d}). In order to extrapolate from our grids to estimate the detection rate at high temperature and low $v\sin{i}$ and at low temperature and high $v\sin{i}$, we made the following assumptions about the shape of the two-dimensional detection rate surface: First, we assumed that if all companions are detected at temperature $T=6500$ K and rotation speed $v\sin{i}$, then all companions with the same $v\sin{i}$ and larger temperature will also be detected (lower right points in Figure \ref{fig:detrate_2d}). Likewise, we assume that if no companions are detected at temperature $T$ and rotation speed $v\sin{i} = 50\ \mathrm{km\ s}^{-1}$, then no companions will be detected at the same temperature and faster rotation speeds (upper left points in Figure \ref{fig:detrate_2d}). Figure \ref{fig:detrate_2d} shows a clear diagonal dividing line between hot, slow rotators that are always detected and cool, fast rotators that never are. Additionally the figure shows that very fast rotators are never detected, regardless of their temperature, because the signal is completely removed when we high-pass filter the data (see Section \ref{sec:companions}).

\begin{figure}
\includegraphics[width=\columnwidth]{HIP_24244_20130919.pdf}
\caption{Detection rate as a function of companion temperature and $v\sin{i}$ for HIP 24244. All companions that are shaded yellow are detectable, while companions in the purple region are never detectable. The grids of squares in the lower left and upper right show the low temperature and high temperature grid points we used in the sensitivity analysis. The remaining squares come from assumptions about the shape of the detection rate surface and allow us to fully interpolate (see text for details).}
\label{fig:detrate_2d}
\end{figure}

\subsection{Marginalization}
By sampling a suitable distribution of $v\sin{i}$ values for a star of each temperature, we marginalize over the rotation speed:

\begin{equation}
Q(T) = \sum_k Q(T, v_k) v_k 
\end{equation} 
where $Q(T, v\sin{i})$ is the surface plotted in Figure \ref{fig:detrate_2d} and $v_k$ are the samples from the distribution of $v\sin{i}$. For $T < 6000$ K, we sample $v\sin{i}$ using the gyrochronology relation given in \citet{Barnes2010b}:

\begin{equation}
\frac{k_Ct}{\tau} = \ln\left ( \frac{P}{P_0} \right ) \frac{k_Ik_C}{2\tau^2} (P^2 - P_0^2)
\label{eqn:gyro}
\end{equation}
In Equation \ref{eqn:gyro}, $k_C$ and $k_I$ are constants fit to data with known ages and rotation periods, $P$ and $P_0$ are respectively the current and zero-age main sequence (ZAMS) rotation periods, $\tau$ is the convective turnover time scale and $t$ is the current age of the star. We use the same values that \cite{Barnes2010b} use for the constants:

\begin{itemize}
\item $k_C = 0.646$ day/Myr
\item $k_I = 452$ Myr/day
\end{itemize}
We estimate the convective timescale ($\tau$) by interpolating Table 1 of \citet{Barnes2010a}. We then randomly sample the system age $t$ from its probability distribution function (see Section \ref{sec:sp} and Table \ref{tab:sample}). Young stars have rotation periods in the range of 0.2 to 10 days \citep{Bouvier2014}, so we randomly choose an initial rotation period $P_0$ from a log-uniform distribution in this range for each age sample. Equation \ref{eqn:gyro} then gives a current rotation period for each sample, which we convert to an equatorial velocity with the stellar radius $R$. We estimate the stellar radius by interpolating Dartmouth pre main sequence isochrones [CITE] using the companion temperature and system age. We finally convert to projected velocity $v\sin{i}$ by randomly sampling a uniform distribution for the inclination $\sin{i}$.

The gyrochronology relations are invalid for stars with $T \gtrsim 6250\ K$, the canonical limit at which the convective zone is too small to efficiently remove angular momentum to the stellar wind and spin down the star \citep{Pinsonneault2001}. \citet{Zorec2012} fit maxwellian distributions to the equatorial velocity of A- and B-type stars in several mass bins. For $T > 7000\ K$, we linearly interpolate the fit parameters as a function of mass and sample the resulting maxwellian probability density function.

Typical velocities from the gyrochronology relationships are $10-20 \mathrm{km\ s}^{-1}$, while the maxwellian velocity distributions have typical velocities $\sim 100 \mathrm{km\ s}^{-1}$. We transition between the two regimes for temperatures in the range $6000\ K < T < 7000\ K$ by first estimating the equatorial velocities from the gyrochronology relationship (Equation \ref{eqn:gyro}) at $T=6000\ K$. We then fit the velocities to a maxwellian distribution, and add the result to the tabulated parameters from \citep{Zorec2012}. With the extended table, we treat stars in the transition range the same way we treat hot stars.

We show the marginalized detection rate and mean value of $v\sin{i}$ as a function of temperature in Figure \ref{fig:marginalized}. Both the detection rate and the average $v\sin{i}$ are smoothly varying, and show the expected behaviour with temperature. The detection rate falls with hotter temperatures because the companions are expected to be fast rotators, which are more difficult to detect.

\begin{figure}
\includegraphics[width=\columnwidth]{HIP_24244_20130919_marginalized.pdf}
\caption{Marginalized detection rate for the same star as shown in Figure \ref{fig:detrate_2d}. The fall in detection rate towards hotter stars is caused by the increase in typical rotational speeds.}
\label{fig:marginalized}
\end{figure}

\subsection{Conversion to Mass Ratio}

The result of the previous analysis is a series of estimates for the detection rate as a function of companion temperature for each observation of each star. We convert companion temperature to mass by interpolating Table 5 of \citet{Pecaut2013}. Next, we estimate the primary mass for each star as the median of the mass samples developed in Section \ref{sec:sp}. We then convert each detection rate curve to be a function of mass ratio ($Q_j(q)$, where $j$ denotes the $j$th star in the sample), and linearly interpolate onto a grid in mass ratio from $0 < q_i < 1$. Finally, we combine the detection rate curves for each star \emph{with no companion detection in our data} into an estimate of the survey-wide completeness by taking the average of the detection rate for all stars:

\begin{equation}
Q(q_i) = \frac{1}{N_i} \sum_j Q_j(q_i)
\label{eqn:completeness}
\end{equation}
In the equation above, $N_i$ is the number of sample stars that contain an estimate for $Q(q_i)$ without extrapolating. For $q_i \sim 0.2$, $N_i$ is near the total sample size. However, $N_i$ falls for both low and high q, since a $3000\ K$/$12000\ K$ companion has a mass ratio $q = 0.08/2.0$ for an A9V primary, but $q = 0.007/0.19$ for a B0V primary. Our sensitivity analysis therefore does not sample large mass ratios around the very early-type primary stars in the sample, and does not sample very low mass ratios around late-type primary stars.

\begin{figure}
\includegraphics[width=\columnwidth]{SurveyCompleteness.pdf}
\caption{Survey completeness as a function of mass ratio ($q$).}
\label{fig:completeness}
\end{figure}

We show the total survey completeness in Figure \ref{fig:completeness}. The completeness falls very rapidly towards low mass ratios, although we are still $\sim 70\%$ complete at $q = 0.1$. The slow fall-off towards large mass ratios is caused by the same effect as in Figure \ref{fig:marginalized}: large mass ratio companions tend to be rapid rotators that are more difficult to detect.

\section{Mass-Ratio Distribution}
\label{sec:mrd}

We are now finally in a position to estimate the mass-ratio distribution for our sample. We estimate the mass for each detected companion star by sampling the temperatures given in Table \ref{tab:companions} as a gaussian, and converting each temperature sample into a mass sample. We do the conversion to mass both by interpolating Table 5 from \citet{Pecaut2013}, and by interpolating from temperature and system age (see Section \ref{sec:sp}) to mass with Dartmouth isochrones [CITE]. Both methods give similar results in most cases, although the isochrone masses are significantly higher for very young systems, so we adopt the isochronal values in the analysis. We sample the mass ratio of the system by dividing the companion mass samples by samples of the primary mass (Section \ref{sec:sp}). We denote the $n$th mass ratio sample for the $k$th star as $q_k^{(n)}$, and denote the number of these samples as $N_k$.

We do not use systems with more than one companion, unless the wider companion is separated by $ > 10''$ from the primary star. We mark the 48 companions we use in the mass ratio analysis with the fourth column of Table \ref{tab:companions}. Many of the companions we use in the analysis only have one detection in our data; 12/22 of these are previously known companions and so don't need follow-up to confirm. The remaining 10 are new and unconfirmed detections; these all have very strong CCF signals and are likely to be confirmed with follow-up spectroscopy or imaging. Their inclusion does not significantly change the results.

\subsection{Fitting  Methodology}

We use the methodology developed in \citet{Foreman2014} to perform bayesian inference on the shape and form of the mass ratio distribution. The log-likelihood function in this formalism is derived from modeling the survey as a draw from the inhomogeneous Poisson process with rate density $\Gamma \equiv KQ(q)P(q)$:

\begin{multline}
\ln{\mathcal{L}(\{\vec{x_k}\}| \vec{\theta})} = -K \int_0^1 Q(q)P(q|\vec{\theta})dq + \\ \sum_{k=1}^K \ln{\frac{K}{N_k} \sum_{n=1}^{N_k} Q(q_k^{(n)}) P(q_k^{(n)}|\vec{\theta})}
\label{eqn:money}
\end{multline}

In the above equation, $\{\vec{x_k}\}$ denotes the full spectrum of star $k$, and $\vec{\theta}$ denotes the parameters for the model we are fitting. $K=360$ is the total sample size for our survey, $Q(q)$ is the completeness function shown in Figure \ref{fig:completeness}, and $P(q|\vec{\theta})$ is the likelihood function for the mass ratio given the model parameters. We fit the data to three distinct distributions: a histogram ($P_1$), a beta distribution ($P_2$), and a power law ($P_3$):

\begin{align}
 P_1(q|\vec{\theta}) &= \begin{cases}
      \hfill \theta_1 \hfill & q \in \Delta_1 \\
      \hfill \theta_2 \hfill & q \in \Delta_2 \\
      \hfill \ldots \\
      \hfill \theta_7 \hfill & q \in \Delta_7
     \end{cases} \label{eqn:P1} \\
 %P_2(q|\vec{\theta}) &= \frac{A}{\sqrt{2\pi\sigma^2}} e^{-\frac{(q - \mu)^2}{2\sigma^2}} \label{eqn:P2} \\
 P_2(q|\vec{\theta}) &= \frac{\Gamma(a+b)}{\Gamma(a)\Gamma(b)} q^{a-1}(1-q)^{b-1} \\
 P_3(q|\vec{\theta}) &= (1-\gamma)q^{-\gamma} \label{eqn:P3}
\end{align}
%The constant $A$ in the gaussian distribution is a renormalization factor such that the distribution is only defined from $0 < q < 1$:

%\begin{equation}
%A = \frac{2}{\erf\left(\frac{\mu}{\sigma \sqrt{2}}\right) - \erf\left(\frac{\mu - 1}{\sigma \sqrt{2}}\right)}
%\end{equation}

We fit all distributions via Importance Nested Sampling with the MultiNest code \citep{multinest}. Following \citet{Foreman2014}, we apply a smoothing prior on the parameters $\vec{\theta}$ for the histogram model:

\begin{align}
P(\vec{\theta}| \alpha, m, \tau, \epsilon) &= \mathcal{N}(\vec{\theta} | m, K(\{\Delta_j\}, \alpha, \tau, \epsilon)) \\
K_{ij} &= \sqrt{\left[\alpha \exp{\left(-\frac{(\Delta_i - \Delta_j)^2}{2\tau^2}\right)}\right]^2 + \epsilon^2 \delta_{ij}}
\end{align}
The smoothing prior is an 7-dimensional gaussian with mean $m$ and covariance matrix $K_{ij}$, and encodes our belief that the mass ratio distribution is a smoothly varying function while leaving enough flexibility to let the data drive the shape of the function. Since we have introduced three new hyperparameters ($a, m, \tau, \epsilon$), we must apply a prior to them and marginalize over them when estimating the bin heights. We choose log-uniform priors for $a, \tau$, and $\epsilon$, and a uniform prior for the mean $m$. The full posterior probability distribution for the histogram model is:

\begin{equation}
P_1(\vec{\theta} | \{\vec{x_k}\}) \propto \mathcal{L}_1(\{\vec{x_k}\}| \vec{\theta}) P(\vec{\theta}| \alpha, m, \tau, \epsilon) P(\alpha, m, \tau, \epsilon)
\end{equation}

The beta function fit only has two parameters ($a, b$), and was chosen because it has a similar shape to the histogram resulting from the first model and is only defined for $0 < q < 1$. We use uniform priors on both $a$ and $b$, resulting in the posterior probability

\begin{equation}
P_2(a, b | \{\vec{x_k}\}) \propto P_2(q|a, b) 
\end{equation}

The power law fit has only one parameter ($\gamma$); we use a uniform prior in the fit.

\subsection{Malmquist Bias Correction}

We are trying to recover the intrinsic distribution from an observed sample, so we must fit the data to the probability distribution function (PDF) for mass ratio, \emph{given that we observed the star}: $P(q|\vec{\theta}, \mathrm{obs})$. In a volume-limited sample, this is equal to $P(q|\vec{\theta})$. However, our sample is magnitude-limited and therefore suffers from Malmquist bias. There is a higher probability of equal-mass binary systems to occur in our survey because they contribute twice the flux and are therefore more likely to fall under the magnitude limit. We can calculate the PDF for mass ratio, given that we observed the system, from Bayes' theorem:

\begin{equation}
P(q|\vec{\theta}, \mathrm{obs}) = \frac{P(\mathrm{obs}|q) P(q|\vec{\theta})}{\int_0^1 P(\mathrm{obs}|q) P(q|\vec{\theta}) dq}
\end{equation}

We already know $P(q|\vec{\theta})$ (Equations \ref{eqn:P1} - \ref{eqn:P3}). We estimate $P(\mathrm{obs}|q)$ by simulating a very large sample of binary stars via these steps:

\begin{enumerate}
\item Draw random primary star masses from the Kroupa IMF \citep{Kroupa2002}
\item Draw a random distance for each star from a disk with infinite extent and scale height of $150$ pc \citep[the approximate scale height of the Milky Way disk for A-type stars,][]{BM1998}. 
\item For each $q$ from 0 to 1, in steps of 0.01:
\begin{enumerate}
  \item Add a companion star to each primary with the appropriate mass to make a binary system with mass ratio $q$.
  \item Calculate the combined absolute V-magnitude by interpolating Table 5 of \citet{Pecaut2013}.
  \item Calculate apparent magnitude $V$ from the absolute magnitude and distance.
  \item Find fraction of stars ($f(q)$) with apparent $V < 6$
\end{enumerate}
\item Fit the sampled fractions $f(q)$ to a 5th-order polynomial.
\end{enumerate}

With the fitted malmquist-correction polynomial, we then substitute $P(q|\vec{\theta}, \mathrm{obs})$ everywhere $P(q|\vec{\theta})$ appears in Equation \ref{eqn:money}.

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 500 samples of the beta distribution fit to show the spread allowed by the data. The best-fit power law is plotted with a green dot-dashed line, as well as the mass ratio distribution expected from random pairing from the Kroupa initial mass function (IMF) in yellow.

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_total.pdf}
\caption{Mass ratio distribution for our sample. The data was fit to a histogram, a beta distribution, and a power law. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the beta function fit is shown with 500 samples from the posterior probability distribution for the parameters. We also show the best-fit power law and the mass ratio distribution resulting from random pairing of the Kroupa initial mass function.}
\label{fig:mrd}
\end{figure}

\section{Discussion}
\label{sec:discussion}

Previous measurements of the mass-ratio distribution find that the data is well fit by a power law. \citet{Kouwenhoven2007} compiled spectroscopic, imaging, and astrometric observations of binary stars with intermediate-mass primaries in the Scorpius OB association, and derived a power law index of  $-0.45 \pm 0.15$. More recently \citet{DeRosa2014} performed an adaptive optics and common proper motion search for companions to field A-type stars. They found that the distribution for companions on wide ($a > 125\ AU$) orbits has a very steep power law index of $-2.3^{+1.0}_{-0.9}$, while the distribution for close ($30\ AU < a < 125\ AU$) companions is consistent with flat.

\subsection{Model Comparison}

The most striking feature of the mass ratio distribution shown in Figure \ref{fig:mrd} is the turnover at intermediate q, which the beta distribution fit finds occurs at $q = 0.30^{+0.03}_{-0.04}$. Although the power law fit and the Kroupa IMF are visually very poor fits to the data, we formally compare the models to ensure that the different form is not an artifact of binning or simple noise. We make the comparison using the posterior odds:

\begin{equation}
\mathrm{Odds} = \frac{\int_{\vec{\theta_1}} P_1(q|\vec{\theta_1}) P(\vec{\theta_1})}{\int_{\vec{\theta_2}} P_2(q|\vec{\theta_2}  P(\vec{\theta_2})} \equiv \frac{Z_1}{Z_2}
\end{equation}

The integrals are estimated as part of the nested sampling algorithm in the MultiNest code. The odds ratio comparing the beta distribution to power law fit is $Z_\mathrm{beta} / Z_\mathrm{power} = 81 \pm 1.3 \e{8}$ indicating a very strong preference for the beta distribution model. We also compare to the mass ratio distribution expected for random pairing from the Kroupa IMF. In this case, there are no free parameters so the evidence integral just becomes the likelihood function (Equation \ref{eqn:money}). The corresponding odds ratio is  $Z_\mathrm{beta} / Z_\mathrm{kroupa} = 325 \pm 3.6 \e{18}$, again indicating a very strong preference for the beta distribution fit. The extreme unlikeliness of the Kroupa IMF model also indicates that our sample is not significantly biased by foreground or background contaminants.

\subsection{Comparison to Previous Results}

Our mass ratio distribution appears to be in tension with the results of the VAST survey \citep{DeRosa2014}, which finds a nearly flat distribution for close companions. However, their subsample of close companions only includes 18 binaries, so it is possible that the different forms are just a result of small number statistics. To assess the degree of tension, we use the Anderson-Darling  test \citep{Anderson1954} to find the probability that both their close companion subsample and our companions are drawn from the same parent distribution. We only use companions from this work with mean $\overline{q} > 0.15$, since the survey completeness is roughly constant in that range and because the VAST survey subsample makes the same cut. To account for measurement uncertainties in the mass ratios, we choose from the mass ratio samples ($q_k^{(n)}$, see Section \ref{sec:mrd}) many times and compute the Anderson-Darling test statistic each time. Since \citet{DeRosa2014} do not quote uncertainties, we assume uncertainties of $\sigma_q = 0.05$ for all of their measurements. The result is $p = 0.02^{+0.02}_{-0.01}$, indicating that it is unlikely that both samples come from the same parent distribution.

The differing results indicate that the mass ratio distribution changes inside $\sim 30-50$ AU, where \citet{DeRosa2014} has low completeness. This intrepretation implies that the mass ratio distribution is a very steep power law at wide ($a \gtrsim 150$ AU) separations, flattens out at intermediate separations ($50\ \mathrm{AU} \lesssim a \lesssim 150\ \mathrm{AU}$), and begins to show a preference for $q \sim 0.3$ at close separations ($a \lesssim 50$ AU).

%\subsection{Theoretical Implications}

%TODO: Discuss ways forward. Compile table giving the estimated maximum separation from the RVs. Promote an imaging survey to follow these up. Promote RV monitoring of the detections to get dynamical mass ratios and, more importantly, separations.

%TODO: Discuss physical mechanism to generate a preferrence for $q = 0.3$. Saying something about when disk fragments (how much time is left for the companion to accrete)? Saying something about migration in the early disk? Perhaps the extreme mass-ratio systems migrate inwards as they accrete material. If they continue migrating long enough to get much over $q \sim 0.3$, they end up falling into the star?

\subsection{Supplemental Data}

In the effort of open and reproducible research, we have made several data products freely available to the community. The samples of primary and companion mass and system age are available at \url{https://zenodo.org/record/46340}. The files generated by MultiNest, which include samples from the posterior distributions described in Section \ref{sec:mrd}, are available at the same url. We additionally provide a series of python libraries and jupyter notebooks with the computer code we used for the analysis on github: \url{https://github.com/kgullikson88/BinaryInference}.

\section*{Acknowledgements}
This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and of Astropy, a community-developed core Python package for Astronomy (Astropy Collaboration, 2013).
It was supported by a start-up grant to Adam Kraus as well as a University of Texas Continuing Fellowship to Kevin Gullikson.

This work used the Immersion Grating Infrared Spectrograph (IGRINS) that was developed under a collaboration between the University of Texas at Austin and the Korea Astronomy and Space Science Institute (KASI) with the financial support of the US National Science Foundation under grant AST-1229522, of the University of Texas at Austin, and of the Korean GMT Project of KASI.

The Hobby-Eberly Telescope (HET) is a joint project of the University of Texas at Austin, the Pennsylvania State University, Stanford University, Ludwig-Maximilians-Universit\"at M\"unchen, and Georg-August-Universit\"at G\"ottingen. The HET is named in honor of its principal benefactors, William P. Hobby and Robert E. Eberly.

Based on observations at Cerro Tololo Inter-American Observatory, National Optical Astronomy Observatory (NOAO Prop. IDs: 13A-0139, 13B-0112, 2014A-0260, 14A-0260, 15A-0245; PI: Kevin Gullikson), which is operated by the Association of Universities for Research in Astronomy (AURA) under a cooperative agreement with the National Science Foundation.

\LongTables
\begin{deluxetable*}{|l|lrrrrllllll|}
\tabletypesize{\scriptsize}
\tablewidth{0pt}
\tablecaption{Sample Properties}
\tablehead{
      \colhead{} & \colhead{} & \colhead{} & \colhead{} &  \colhead{}  & \colhead{parallax} & \colhead{$T_\mathrm{eff}$} & \colhead{$\log{g}$} & \colhead{Mass} & \colhead{Age} & \colhead{} \\
       \colhead{star} & \colhead{SpT} & \colhead{RA} & \colhead{DEC} &  \colhead{V} &   \colhead{
       (mas)} & \colhead{(K)} & \colhead{(cgs)} & \colhead{($M_{\odot}$)} & \colhead{(Myr)} & \colhead{Ref}
       }

\startdata

     HIP 813 &     B9Vn &    00:10:02.20 &   +11:08:44.93 &   5.537 &     10.68 &   $12516 \pm 426$ &  $4.3 \pm 0.14$ &  $3.1^{+0.18}_{-0.17}$ &      $85^{+56}_{-52}$ &       1 \\
    HIP 1191 &    B8.5V &    00:14:54.52 &   -09:34:10.45 &   5.757 &      9.63 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.41}_{-0.37}$ &      $23^{+63}_{-16}$ &       2 \\
    HIP 1366 &      A2V &    00:17:05.50 &   +38:40:53.89 &   4.610 &     10.56 &    $9371 \pm 319$ &  $4.0 \pm 0.14$ &  $2.2^{+0.18}_{-0.16}$ &    $464^{+83}_{-119}$ &       1 \\
    HIP 1647 &      B9V &    00:20:39.04 &   -69:37:29.68 &   5.498 &     10.25 &   $11393 \pm 387$ &  $4.0 \pm 0.14$ &  $2.8^{+0.21}_{-0.18}$ &     $206^{+48}_{-79}$ &       1 \\
    HIP 2381 &      A3V &    00:30:22.65 &   -23:47:15.65 &   5.190 &     18.83 &    $8364 \pm 284$ &  $4.0 \pm 0.14$ &  $1.9^{+0.16}_{-0.13}$ &   $715^{+135}_{-183}$ &       1 \\
    HIP 2505 &     B8Vn &    00:31:46.36 &   +54:31:20.23 &   4.732 &      8.64 &  $12000 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.9^{+0.45}_{-0.40}$ &     $58^{+104}_{-48}$ &       2 \\
    HIP 2548 &    B9.5V &    00:32:23.78 &   +06:57:19.66 &   5.698 &     12.35 &   $11864 \pm 403$ &  $4.4 \pm 0.14$ &  $2.8^{+0.16}_{-0.15}$ &      $77^{+69}_{-51}$ &       1 \\
    HIP 2912 &      B5V &    00:36:52.85 &   +33:43:09.64 &   4.360 &      5.45 &  $16000 \pm 1000$ &  $4.0 \pm 0.25$ &  $4.7^{+0.61}_{-0.54}$ &      $26^{+32}_{-18}$ &       2 \\
    HIP 3300 &      B2V &    00:42:03.90 &   +50:30:45.09 &   4.810 &      2.28 &  $18000 \pm 1000$ &  $4.0 \pm 0.25$ &  $5.7^{+0.67}_{-0.63}$ &      $19^{+19}_{-13}$ &       2 \\
    HIP 3478 &      B5V &    00:44:26.19 &   +47:51:50.34 &   5.646 &      5.23 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.60}_{-0.57}$ &       $11^{+14}_{-5}$ &       2 \\
    HIP 3741 &      B9V &    00:48:01.06 &   -21:43:20.99 &   5.570 &      9.51 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.47}_{-0.42}$ &    $282^{+82}_{-115}$ &       2 \\
    HIP 3881 &      B5V &    00:49:48.85 &   +41:04:44.08 &   4.530 &      5.28 &  $16000 \pm 1000$ &  $4.2 \pm 0.25$ &  $4.5^{+0.58}_{-0.52}$ &      $17^{+28}_{-11}$ &       2 \\
    HIP 4436 &      A5V &    00:56:45.21 &   +38:29:57.64 &   3.870 &     25.14 &    $8351 \pm 284$ &  $4.0 \pm 0.14$ &  $1.9^{+0.17}_{-0.14}$ &   $733^{+129}_{-158}$ &       1 \\
    HIP 5131 &     A1Vn &    01:05:40.96 &   +21:28:23.45 &   5.317 &     11.86 &   $11956 \pm 406$ &  $4.4 \pm 0.14$ &  $2.8^{+0.16}_{-0.14}$ &      $69^{+65}_{-45}$ &       1 \\
    HIP 5132 &     A0Vn &    01:05:41.71 &   +21:27:55.60 &   5.532 &     11.64 &   $12053 \pm 410$ &  $4.4 \pm 0.14$ &  $2.9^{+0.16}_{-0.14}$ &      $64^{+61}_{-42}$ &       1 \\
    HIP 5310 &      A3V &    01:07:57.16 &   +20:44:20.83 &   5.569 &     21.14 &    $8611 \pm 293$ &  $4.4 \pm 0.14$ &  $1.8^{+0.10}_{-0.08}$ &   $307^{+228}_{-196}$ &       1 \\
    HIP 5348 &  B6V+B9V &    01:08:23.08 &   -55:14:44.73 &   3.967 &     10.92 &   $13776 \pm 468$ &  $4.2 \pm 0.14$ &  $3.6^{+0.23}_{-0.21}$ &      $67^{+38}_{-39}$ &       1 \\
    HIP 5361 &      B8V &    01:08:33.47 &   +58:15:48.41 &   5.773 &      5.41 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.49}_{-0.43}$ &      $18^{+36}_{-11}$ &       2 \\
    HIP 5518 &    A0Vnn &    01:10:39.32 &   +68:46:43.04 &   5.318 &     11.78 &   $11894 \pm 404$ &  $4.1 \pm 0.14$ &  $3.0^{+0.21}_{-0.19}$ &     $172^{+40}_{-72}$ &       1 \\
    HIP 5626 &      A3V &    01:12:16.82 &   +79:40:26.27 &   5.600 &     12.07 &   $10342 \pm 352$ &  $4.2 \pm 0.14$ &  $2.4^{+0.15}_{-0.14}$ &   $218^{+102}_{-127}$ &       1 \\
    HIP 7345 &      A1V &    01:34:37.78 &   -15:40:34.90 &   5.619 &     16.84 &   $10007 \pm 340$ &  $4.4 \pm 0.14$ &  $2.2^{+0.12}_{-0.11}$ &   $156^{+130}_{-100}$ &       1 \\
    HIP 8016 &      B9V &    01:42:55.86 &   +70:37:21.09 &   5.177 &     11.75 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.21}$ &     $67^{+188}_{-57}$ &       2 \\
    HIP 8704 &    B1.5V &    01:51:59.32 &   +55:08:50.58 &   5.520 &      2.52 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.74}_{-0.70}$ &         $8^{+6}_{-3}$ &       2 \\
      HR 545 &      B9V &    01:53:31.77 &   +19:17:46.27 &   4.700 &   \nodata &   $10000 \pm 500$ &  $4.5 \pm 0.25$ &  $2.2^{+0.21}_{-0.20}$ &     $34^{+106}_{-27}$ &       2 \\
    HIP 9312 &     A0Vn &    01:59:38.04 &   +64:37:17.76 &   5.283 &     13.28 &   $11913 \pm 405$ &  $4.1 \pm 0.14$ &  $2.8^{+0.13}_{-0.11}$ &      $77^{+65}_{-50}$ &       1 \\
    HIP 9564 &     A1Vn &    02:02:52.48 &   +64:54:05.27 &   5.999 &     11.92 &   $11266 \pm 383$ &  $4.3 \pm 0.14$ &  $2.7^{+0.15}_{-0.12}$ &     $149^{+74}_{-88}$ &       1 \\
      HR 604 &      B8V &    02:03:54.72 &   +42:19:51.41 &   5.820 &   \nodata &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.46}_{-0.42}$ &    $281^{+82}_{-117}$ &       2 \\
   HIP 10320 &      B9V &    02:12:54.47 &   -30:43:25.77 &   5.261 &     10.18 &   $11745 \pm 399$ &  $3.7 \pm 0.14$ &  $3.1^{+0.23}_{-0.19}$ &     $205^{+22}_{-43}$ &       1 \\
   HIP 10670 &    A1Vnn &    02:17:18.87 &   +33:50:49.90 &   4.000 &     29.04 &   $10772 \pm 366$ &  $4.2 \pm 0.14$ &  $2.6^{+0.16}_{-0.13}$ &    $230^{+63}_{-109}$ &       1 \\
   HIP 10732 &     A1Vn &    02:18:07.54 &   +19:54:04.19 &   5.575 &      7.29 &   $9500 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.1^{+0.37}_{-0.31}$ &    $107^{+286}_{-93}$ &       2 \\
   HIP 11345 &      A0V &    02:25:57.01 &   -12:17:25.71 &   4.869 &      7.15 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.22}$ &     $72^{+206}_{-60}$ &       2 \\
   HIP 12332 &      A7V &    02:38:48.99 &   +21:57:41.06 &   5.454 &      9.68 &    $8000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.0^{+0.40}_{-0.33}$ &   $621^{+269}_{-268}$ &       2 \\
   HIP 12706 &     A2Vn &    02:43:18.04 &   +03:14:08.94 &   3.470 &     40.97 &    $8551 \pm 291$ &  $4.3 \pm 0.14$ &  $1.9^{+0.15}_{-0.13}$ &   $647^{+104}_{-184}$ &       1 \\
   HIP 12719 &      B3V &    02:43:27.11 &   +27:42:25.72 &   4.670 &      9.51 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.62}_{-0.57}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 12803 &     B9Vn &    02:44:32.97 &   +15:18:42.71 &   5.776 &      5.49 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.39}_{-0.36}$ &      $25^{+63}_{-17}$ &       2 \\
   HIP 13165 &      B6V &    02:49:17.56 &   +17:27:51.52 &   5.314 &      4.18 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.4^{+0.55}_{-0.51}$ &       $13^{+22}_{-7}$ &       2 \\
   HIP 13202 &      A0V &    02:49:54.18 &   -27:56:31.14 &   5.389 &      7.11 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.44}_{-0.38}$ &   $401^{+138}_{-170}$ &       2 \\
   HIP 13209 &     B8Vn &    02:49:59.03 &   +27:15:37.83 &   3.606 &     19.69 &   $13316 \pm 453$ &  $4.1 \pm 0.14$ &  $3.3^{+0.15}_{-0.13}$ &      $45^{+43}_{-30}$ &       1 \\
   HIP 13327 &      B7V &    02:51:29.59 &   +15:04:55.45 &   5.514 &      6.60 &  $14000 \pm 1000$ &  $4.0 \pm 0.25$ &  $3.8^{+0.55}_{-0.46}$ &      $36^{+57}_{-27}$ &       2 \\
   HIP 13717 &      A3V &    02:56:37.42 &   -03:42:44.35 &   5.160 &     17.49 &    $8612 \pm 293$ &  $4.0 \pm 0.14$ &  $1.9^{+0.13}_{-0.11}$ &   $571^{+138}_{-255}$ &       1 \\
   HIP 13879 &     A2Vn &    02:58:45.67 &   +39:39:45.81 &   4.700 &     10.53 &    $9298 \pm 316$ &  $3.5 \pm 0.14$ &  $2.1^{+0.12}_{-0.09}$ &   $340^{+141}_{-193}$ &       1 \\
   HIP 14043 &      B7V &    03:00:52.21 &   +52:21:06.22 &   5.253 &      7.11 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.48}_{-0.45}$ &      $17^{+33}_{-10}$ &       2 \\
   HIP 14143 &      B7V &    03:02:22.52 &   +04:21:10.39 &   5.607 &      5.93 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.50}_{-0.45}$ &      $17^{+36}_{-11}$ &       2 \\
   HIP 14293 &      A5V &    03:04:16.52 &   -07:36:03.08 &   5.300 &     24.06 &    $8039 \pm 273$ &  $4.2 \pm 0.14$ &  $1.8^{+0.14}_{-0.10}$ &   $782^{+166}_{-291}$ &       1 \\
   HIP 14576 &      B8V &    03:08:10.13 &   +40:57:20.33 &   2.120 &     36.27 &   $13070 \pm 444$ &  $4.0 \pm 0.14$ &  $3.6^{+0.29}_{-0.24}$ &     $139^{+16}_{-25}$ &       1 \\
   HIP 14764 &      B8V &    03:10:38.79 &   +11:52:21.44 &   5.965 &      7.33 &  $12000 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.9^{+0.47}_{-0.40}$ &     $54^{+103}_{-44}$ &       2 \\
   HIP 14862 &    A2Vnn &    03:11:56.27 &   +74:23:37.17 &   4.840 &     19.72 &   $8875 \pm 1000$ &  $4.2 \pm 0.25$ &  $1.8^{+0.32}_{-0.29}$ &     $71^{+317}_{-60}$ &       2 \\
   HIP 15110 &      A1V &    03:14:54.10 &   +21:02:40.01 &   4.880 &     12.44 &    $9902 \pm 337$ &  $4.0 \pm 0.14$ &  $2.4^{+0.18}_{-0.15}$ &    $375^{+52}_{-104}$ &       1 \\
   HIP 15338 &      B8V &    03:17:47.35 &   +44:01:30.08 &   5.478 &      4.46 &  $12000 \pm 1000$ &  $3.5 \pm 0.25$ &  $3.5^{+0.65}_{-0.59}$ &     $155^{+68}_{-59}$ &       2 \\
   HIP 15404 &      B3V &    03:18:37.74 &   +50:13:19.83 &   5.158 &      5.12 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.5^{+0.68}_{-0.64}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 15444 &      B5V &    03:19:07.64 &   +50:05:41.88 &   5.036 &      5.82 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.5^{+0.60}_{-0.60}$ &       $11^{+14}_{-6}$ &       2 \\
   HIP 16147 &      B5V &    03:28:03.07 &   +49:03:46.32 &   4.970 &      6.22 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.59}_{-0.58}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 16210 &     B6Vn &    03:28:52.33 &   +49:50:54.17 &   5.578 &      6.22 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 16244 &      B3V &    03:29:22.05 &   +49:30:32.21 &   4.678 &      6.05 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.61}_{-0.57}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 16285 &      A5V &    03:29:55.15 &   -42:38:03.32 &   5.768 &     15.28 &    $7884 \pm 268$ &  $3.9 \pm 0.14$ &  $1.7^{+0.13}_{-0.09}$ &   $841^{+183}_{-335}$ &       1 \\
   HIP 16322 &     A0Vn &    03:30:24.47 &   +11:20:11.19 &   5.125 &      9.03 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.26}_{-0.22}$ &     $80^{+200}_{-69}$ &       2 \\
   HIP 16340 &      B8V &    03:30:36.95 &   +48:06:12.95 &   5.820 &      3.65 &  $12000 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.9^{+0.48}_{-0.40}$ &     $54^{+107}_{-44}$ &       2 \\
   HIP 16599 &      A3V &    03:33:39.06 &   +54:58:29.49 &   5.981 &     13.44 &    $8383 \pm 285$ &  $4.0 \pm 0.14$ &  $2.0^{+0.19}_{-0.15}$ &   $737^{+102}_{-102}$ &       1 \\
   HIP 16611 &      B9V &    03:33:47.28 &   -21:37:58.38 &   4.300 &     11.12 &   $12514 \pm 425$ &  $4.0 \pm 0.14$ &  $3.3^{+0.24}_{-0.20}$ &     $157^{+23}_{-45}$ &       1 \\
   HIP 17457 &     B7IV &    03:44:30.51 &   -01:09:47.14 &   5.250 &      4.98 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.47}_{-0.44}$ &      $17^{+36}_{-10}$ &       2 \\
   HIP 17527 &      B8V &    03:45:09.74 &   +24:50:21.34 &   5.640 &      7.97 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.47}_{-0.44}$ &      $18^{+34}_{-11}$ &       2 \\
   HIP 17563 &      B3V &    03:45:40.44 &   +06:02:59.97 &   5.332 &      6.11 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.64}_{-0.57}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 18141 &      B8V &    03:52:41.66 &   -05:21:40.54 &   5.476 &      5.57 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.47}_{-0.43}$ &      $18^{+35}_{-12}$ &       2 \\
   HIP 18396 &      B6V &    03:55:58.17 &   +47:52:17.12 &   5.379 &      4.47 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.54}_{-0.50}$ &       $14^{+21}_{-8}$ &       2 \\
   HIP 18724 &      B3V &    04:00:40.82 &   +12:29:25.22 &   3.410 &      6.74 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 18788 &      B5V &    04:01:32.05 &   -01:32:58.78 &   5.280 &      7.88 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.51}_{-0.49}$ &       $13^{+22}_{-7}$ &       2 \\
   HIP 18805 &      B5V &    04:01:46.14 &   +09:59:52.84 &   5.676 &      5.71 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.60}_{-0.57}$ &       $10^{+14}_{-5}$ &       2 \\
   HIP 19799 &     B9Vn &    04:14:36.23 &   +10:00:41.05 &   5.208 &      8.34 &   $10000 \pm 500$ &  $4.5 \pm 0.25$ &  $2.2^{+0.21}_{-0.19}$ &     $34^{+116}_{-26}$ &       2 \\
   HIP 19949 &     A2Vn &    04:16:43.09 &   +53:36:42.47 &   5.200 &      9.98 &    $9825 \pm 334$ &  $3.8 \pm 0.14$ &  $2.2^{+0.11}_{-0.10}$ &   $197^{+140}_{-122}$ &       1 \\
   HIP 19968 &      B7V &    04:16:53.56 &   +61:50:59.97 &   5.700 &      7.31 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.53}_{-0.51}$ &       $13^{+21}_{-8}$ &       2 \\
   HIP 20264 &      A0V &    04:20:39.01 &   -20:38:22.64 &   5.380 &      6.86 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.21}$ &     $77^{+207}_{-65}$ &       2 \\
   HIP 20380 &      A3V &    04:21:51.81 &   +56:30:22.74 &   5.920 &     10.31 &    $8738 \pm 297$ &  $3.9 \pm 0.14$ &  $1.9^{+0.10}_{-0.08}$ &   $338^{+206}_{-209}$ &       1 \\
   HIP 20430 &    B9Vnn &    04:22:34.94 &   +25:37:45.54 &   5.376 &     11.20 &   $11981 \pm 407$ &  $4.2 \pm 0.14$ &  $3.2^{+0.28}_{-0.23}$ &     $195^{+23}_{-27}$ &       1 \\
   HIP 20507 &      A2V &    04:23:40.85 &   -03:44:43.68 &   5.171 &     15.60 &    $8793 \pm 299$ &  $4.0 \pm 0.14$ &  $1.9^{+0.10}_{-0.08}$ &   $340^{+200}_{-210}$ &       1 \\
   HIP 20579 &      B8V &    04:24:29.16 &   +34:07:50.73 &   5.722 &      7.28 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.46}_{-0.46}$ &      $17^{+32}_{-10}$ &       2 \\
   HIP 20789 &      B7V &    04:27:17.45 &   +22:59:46.80 &   5.515 &      8.63 &  $14000 \pm 1000$ &  $4.0 \pm 0.25$ &  $3.7^{+0.53}_{-0.47}$ &      $38^{+55}_{-28}$ &       2 \\
   HIP 21589 &      A6V &    04:38:09.46 &   +12:30:39.01 &   4.270 &     21.24 &    $8591 \pm 292$ &  $3.9 \pm 0.14$ &  $1.8^{+0.09}_{-0.07}$ &   $300^{+231}_{-193}$ &       1 \\
   HIP 21683 &     A5Vn &    04:39:16.50 &   +15:55:04.70 &   4.675 &     20.97 &    $8165 \pm 278$ &  $4.0 \pm 0.14$ &  $1.7^{+0.07}_{-0.06}$ &   $318^{+266}_{-211}$ &       1 \\
   HIP 21819 &      A2V &    04:41:19.76 &   +28:36:53.98 &   5.726 &      8.55 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.41}_{-0.37}$ &   $411^{+135}_{-158}$ &       2 \\
   HIP 21928 &     A1Vn &    04:42:54.33 &   +43:21:54.53 &   5.301 &     13.52 &   $10734 \pm 365$ &  $4.0 \pm 0.14$ &  $2.6^{+0.16}_{-0.13}$ &    $242^{+59}_{-108}$ &       1 \\
   HIP 22028 &      A1V &    04:44:07.98 &   -18:39:59.71 &   5.527 &     10.66 &   $10118 \pm 344$ &  $4.0 \pm 0.14$ &  $2.3^{+0.14}_{-0.11}$ &   $246^{+105}_{-135}$ &       1 \\
   HIP 22509 &     A1Vn &    04:50:36.72 &   +08:54:00.65 &   4.350 &     14.53 &    $9784 \pm 333$ &  $2.7 \pm 0.14$ &  $2.2^{+0.11}_{-0.10}$ &   $205^{+141}_{-129}$ &       1 \\
   HIP 22833 &      A3V &    04:54:46.90 &   +11:25:33.63 &   5.186 &     14.29 &    $8278 \pm 281$ &  $3.9 \pm 0.14$ &  $1.8^{+0.12}_{-0.09}$ &   $645^{+165}_{-304}$ &       1 \\
   HIP 22840 &      B5V &    04:54:50.71 &   +00:28:01.81 &   5.975 &      4.50 &  $16000 \pm 1000$ &  $4.0 \pm 0.25$ &  $4.7^{+0.61}_{-0.55}$ &      $27^{+32}_{-18}$ &       2 \\
   HIP 22913 &      B9V &    04:55:50.15 &   +15:02:25.00 &   5.785 &      8.97 &  $12000 \pm 1000$ &  $3.5 \pm 0.25$ &  $3.5^{+0.67}_{-0.57}$ &     $155^{+68}_{-64}$ &       2 \\
   HIP 22958 &      B6V &    04:56:24.19 &   -05:10:16.87 &   5.490 &      4.40 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.52}_{-0.53}$ &       $13^{+20}_{-7}$ &       2 \\
   HIP 23362 &      B9V &    05:01:25.58 &   -20:03:06.91 &   4.894 &     16.48 &   $12450 \pm 423$ &  $4.3 \pm 0.14$ &  $3.2^{+0.21}_{-0.16}$ &     $147^{+28}_{-59}$ &       1 \\
   HIP 23916 &      B8V &    05:08:20.19 &   -08:39:55.17 &   5.780 &      4.79 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 24244 &   B7.5Vn &    05:12:17.90 &   -11:52:09.19 &   4.450 &     14.07 &   $13781 \pm 469$ &  $4.2 \pm 0.14$ &  $3.9^{+0.29}_{-0.23}$ &     $114^{+11}_{-23}$ &       1 \\
   HIP 24327 &      B7V &    05:13:13.88 &   -12:56:28.65 &   4.421 &      4.48 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.47}_{-0.44}$ &      $18^{+35}_{-11}$ &       2 \\
   HIP 24505 &      B9V &    05:15:24.37 &   -26:56:36.63 &   5.040 &     11.73 &   $11748 \pm 399$ &  $4.2 \pm 0.14$ &  $2.9^{+0.16}_{-0.14}$ &     $152^{+53}_{-80}$ &       1 \\
   HIP 24902 &      A3V &    05:20:14.67 &   +41:05:10.35 &   5.468 &     11.77 &    $8275 \pm 281$ &  $4.0 \pm 0.14$ &  $1.8^{+0.09}_{-0.08}$ &   $457^{+261}_{-281}$ &       1 \\
   HIP 25048 &      B3V &    05:21:48.42 &   +41:48:16.46 &   5.200 &      6.16 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.5^{+0.62}_{-0.61}$ &       $11^{+13}_{-5}$ &       2 \\
 ADS 3962 AB &     B1Vn &    05:22:50.30 &   +03:32:52.00 &   4.990 &   \nodata &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 25143 &      A3V &    05:22:50.31 &   +41:01:45.33 &   5.545 &     11.15 &    $8101 \pm 275$ &  $3.8 \pm 0.14$ &  $1.7^{+0.06}_{-0.06}$ &   $231^{+250}_{-156}$ &       1 \\
   HIP 25280 &      A0V &    05:24:28.49 &   -16:58:32.81 &   5.644 &     14.65 &   $10398 \pm 354$ &  $4.4 \pm 0.14$ &  $2.4^{+0.13}_{-0.11}$ &   $199^{+101}_{-115}$ &       1 \\
   HIP 25281 &      B1V &    05:24:28.62 &   -02:23:49.73 &   3.350 &      3.34 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 25555 &   B9.5Vn &    05:27:45.61 &   +15:52:26.58 &   5.512 &      7.69 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.45}_{-0.43}$ &    $283^{+79}_{-104}$ &       2 \\
   HIP 25608 &      A1V &    05:28:15.34 &   -37:13:50.75 &   5.562 &     11.39 &    $9960 \pm 339$ &  $4.1 \pm 0.14$ &  $2.3^{+0.15}_{-0.12}$ &    $311^{+83}_{-149}$ &       1 \\
   HIP 25695 &     B9Vn &    05:29:16.50 &   +25:09:00.78 &   5.480 &      7.67 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.43}_{-0.37}$ &      $24^{+63}_{-17}$ &       2 \\
   HIP 25790 &     A3Vn &    05:30:26.16 &   +15:21:37.61 &   5.940 &     12.89 &    $8397 \pm 286$ &  $3.9 \pm 0.14$ &  $1.8^{+0.08}_{-0.07}$ &   $329^{+250}_{-212}$ &       1 \\
   HIP 25813 &      B5V &    05:30:47.05 &   +05:56:53.29 &   4.200 &     10.77 &   $15603 \pm 531$ &  $4.4 \pm 0.14$ &  $4.7^{+0.35}_{-0.28}$ &       $71^{+8}_{-16}$ &       1 \\
   HIP 26063 &      B1V &    05:33:31.45 &   -01:09:21.87 &   5.340 &      2.22 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 26093 &      B3V &    05:33:54.28 &   +14:18:20.08 &   5.588 &      7.31 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.66}_{-0.64}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 26126 &      A2V &    05:34:16.77 &   +03:46:00.82 &   5.332 &     11.82 &    $9542 \pm 324$ &  $3.9 \pm 0.14$ &  $2.1^{+0.09}_{-0.07}$ &   $177^{+150}_{-115}$ &       1 \\
     HR 1880 &    B0.5V &    05:35:08.48 &   +09:56:06.10 &   6.322 &   \nodata &  \nodata          &  \nodata        &  \nodata                &         \nodata          &       \nodata \\
   HIP 26563 &     A4Vn &    05:38:53.08 &   -07:12:46.18 &   4.800 &     22.42 &    $8416 \pm 286$ &  $4.1 \pm 0.14$ &  $1.8^{+0.08}_{-0.07}$ &   $344^{+256}_{-218}$ &       1 \\
   HIP 27100 &      A7V &    05:44:46.38 &   -65:44:07.90 &   4.360 &     21.80 &    $7828 \pm 266$ &  $3.9 \pm 0.14$ &  $1.8^{+0.18}_{-0.15}$ &   $965^{+153}_{-148}$ &       1 \\
   HIP 27321 &      A6V &    05:47:17.09 &   -51:03:59.44 &   3.860 &     51.44 &    $8300 \pm 282$ &  $4.4 \pm 0.14$ &  $1.8^{+0.11}_{-0.09}$ &   $528^{+235}_{-300}$ &       1 \\
   HIP 27713 &     A2Vn &    05:52:07.73 &   -09:02:30.84 &   5.964 &     10.11 &    $8474 \pm 288$ &  $3.9 \pm 0.14$ &  $1.8^{+0.09}_{-0.07}$ &   $383^{+238}_{-240}$ &       1 \\
   HIP 28691 &      B5V &    06:03:27.37 &   +19:41:26.02 &   5.135 &      4.54 &  $14000 \pm 1000$ &  $4.0 \pm 0.25$ &  $3.8^{+0.54}_{-0.48}$ &      $36^{+55}_{-27}$ &       2 \\
   HIP 28756 &    B2.5V &    06:04:20.27 &   -32:10:20.74 &   5.631 &      3.24 &  $20000 \pm 1000$ &  $4.0 \pm 0.25$ &  $6.9^{+0.75}_{-0.68}$ &       $14^{+13}_{-8}$ &       2 \\
   HIP 28910 &      A0V &    06:06:09.32 &   -14:56:06.92 &   4.669 &     18.88 &   $10453 \pm 355$ &  $4.1 \pm 0.14$ &  $2.4^{+0.16}_{-0.14}$ &    $252^{+80}_{-122}$ &       1 \\
   HIP 29150 &      A0V &    06:08:57.87 &   -22:25:38.68 &   5.482 &     13.24 &   $11283 \pm 384$ &  $4.2 \pm 0.14$ &  $2.6^{+0.12}_{-0.11}$ &     $112^{+82}_{-71}$ &       1 \\
   HIP 29151 &     A3Vn &    06:08:57.90 &   +02:29:58.89 &   5.730 &      4.98 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.43}_{-0.37}$ &   $406^{+141}_{-177}$ &       2 \\
   HIP 29735 &      B9V &    06:15:44.89 &   -13:43:06.29 &   4.998 &      8.00 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.48}_{-0.41}$ &    $281^{+84}_{-105}$ &       2 \\
   HIP 29997 &     A0Vn &    06:18:50.78 &   +69:19:11.23 &   4.762 &     18.64 &   $10834 \pm 368$ &  $4.2 \pm 0.14$ &  $2.5^{+0.12}_{-0.10}$ &     $119^{+95}_{-77}$ &       1 \\
   HIP 30069 &      B9V &    06:19:40.96 &   -34:23:47.73 &   5.750 &      8.00 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.42}_{-0.38}$ &      $25^{+62}_{-18}$ &       2 \\
   HIP 30073 &     B2V  &  06:19:42.7989 &  -07:49:22.473 &   5.246 &      3.74 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.71}_{-0.71}$ &         $8^{+6}_{-3}$ &       2 \\
   HIP 30666 &     A3Vn &    06:26:39.59 &   -01:30:26.41 &   5.874 &     13.80 &   $10520 \pm 358$ &  $4.2 \pm 0.14$ &  $2.6^{+0.20}_{-0.16}$ &     $304^{+37}_{-72}$ &       1 \\
   HIP 30788 &      B4V &    06:28:10.21 &   -32:34:48.25 &   4.480 &      7.70 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.59}_{-0.57}$ &       $11^{+13}_{-6}$ &       2 \\
   HIP 31278 &     B5Vn &    06:33:37.92 &   -01:13:12.55 &   5.083 &      5.89 &  $16000 \pm 1000$ &  $4.0 \pm 0.25$ &  $4.7^{+0.61}_{-0.56}$ &      $26^{+31}_{-18}$ &       2 \\
   HIP 31362 &      B8V &    06:34:35.33 &   -32:42:58.51 &   5.610 &      6.20 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.41}_{-0.37}$ &      $24^{+62}_{-17}$ &       2 \\
   HIP 31434 &    A0Vnn &    06:35:12.06 &   +28:01:20.32 &   5.266 &      8.82 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.7^{+0.46}_{-0.41}$ &    $281^{+83}_{-131}$ &       2 \\
   HIP 32474 &    B9.5V &    06:46:39.02 &   -10:06:26.50 &   5.653 &      6.04 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.47}_{-0.43}$ &    $281^{+85}_{-117}$ &       2 \\
   HIP 32607 &      A8V &    06:48:11.46 &   -61:56:29.00 &   3.300 &     33.78 &    $7770 \pm 264$ &  $3.8 \pm 0.14$ &  $1.6^{+0.07}_{-0.06}$ &   $519^{+344}_{-333}$ &       1 \\
   HIP 33372 &     B8Vn &    06:56:25.83 &   +09:57:23.67 &   5.905 &      7.18 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.47}_{-0.46}$ &      $18^{+37}_{-11}$ &       2 \\
   HIP 33575 &      B2V &    06:58:35.90 &   -25:24:50.97 &   5.577 &      3.78 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.69}_{-0.67}$ &         $8^{+6}_{-3}$ &       2 \\
   HIP 34769 &      A2V &    07:11:51.86 &   -00:29:33.96 &   4.150 &      8.49 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.43}_{-0.38}$ &   $405^{+135}_{-207}$ &       2 \\
   HIP 35180 &      A1V &    07:16:14.55 &   -15:35:08.49 &   5.454 &     12.12 &    $9562 \pm 325$ &  $4.1 \pm 0.14$ &  $2.1^{+0.12}_{-0.09}$ &   $267^{+139}_{-158}$ &       1 \\
   HIP 35341 &     A5Vn &    07:18:02.22 &   +40:53:00.22 &   5.870 &     11.73 &    $8014 \pm 272$ &  $3.9 \pm 0.14$ &  $1.7^{+0.07}_{-0.06}$ &   $357^{+293}_{-237}$ &       1 \\
   HIP 36393 &      A4V &    07:29:20.44 &   +28:07:05.79 &   5.072 &     18.51 &    $9184 \pm 312$ &  $4.1 \pm 0.14$ &  $2.1^{+0.15}_{-0.13}$ &   $431^{+128}_{-200}$ &       1 \\
   HIP 36760 &     A1Vn &    07:33:36.48 &   +15:49:35.98 &   5.269 &      7.70 &    $9000 \pm 500$ &  $3.7 \pm 0.25$ &  $2.1^{+0.40}_{-0.26}$ &   $351^{+175}_{-318}$ &       2 \\
   HIP 36812 &    A0Vnn &    07:34:15.89 &   +03:22:18.19 &   5.830 &      5.67 &   $9500 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.1^{+0.40}_{-0.31}$ &    $108^{+273}_{-95}$ &       2 \\
   HIP 36917 &      B7V &    07:35:22.89 &   -28:22:09.57 &   4.630 &     14.72 &   $13926 \pm 473$ &  $4.3 \pm 0.14$ &  $3.7^{+0.17}_{-0.15}$ &      $62^{+35}_{-37}$ &       1 \\
     HR 2948 &      B6V &    07:38:49.38 &   -26:48:06.49 &   4.500 &   \nodata &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.48}_{-0.44}$ &      $17^{+36}_{-11}$ &       2 \\
   HIP 37297 &    B2.5V &    07:39:27.34 &   -38:18:28.88 &   4.840 &      5.87 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.67}_{-0.62}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 37322 &      B5V &    07:39:43.81 &   -38:08:21.44 &   5.664 &      5.70 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.50}_{-0.50}$ &       $13^{+22}_{-7}$ &       2 \\
   HIP 37450 &      B5V &    07:41:15.81 &   -38:32:00.72 &   5.410 &      5.50 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.4^{+0.53}_{-0.48}$ &       $13^{+22}_{-8}$ &       2 \\
   HIP 38538 &      A3V &    07:53:29.81 &   +26:45:56.82 &   4.977 &     14.66 &    $8551 \pm 291$ &  $4.0 \pm 0.14$ &  $1.9^{+0.15}_{-0.12}$ &   $637^{+111}_{-199}$ &       1 \\
   HIP 38593 &      B2V &    07:54:11.01 &   -35:52:38.23 &   5.462 &      4.81 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.77}_{-0.68}$ &         $8^{+6}_{-3}$ &       2 \\
   HIP 38846 &    B2.5V &    07:56:57.80 &   -43:30:01.48 &   5.340 &      2.08 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.75}_{-0.72}$ &         $8^{+6}_{-3}$ &       2 \\
   HIP 39095 &      A1V &    07:59:52.05 &   -18:23:57.23 &   4.610 &     13.52 &    $8872 \pm 302$ &  $3.4 \pm 0.14$ &  $2.1^{+0.18}_{-0.15}$ &     $593^{+74}_{-85}$ &       1 \\
   HIP 39236 &   B9.5Vn &    08:01:30.29 &   +16:27:19.12 &   5.990 &      6.04 &   $10000 \pm 500$ &  $4.5 \pm 0.25$ &  $2.2^{+0.21}_{-0.19}$ &     $36^{+119}_{-27}$ &       2 \\
   HIP 39567 &      A1V &    08:05:04.49 &   +13:07:05.58 &   5.146 &     15.20 &   $10352 \pm 352$ &  $4.3 \pm 0.14$ &  $2.4^{+0.12}_{-0.10}$ &   $178^{+110}_{-109}$ &       1 \\
   HIP 39847 &      A2V &    08:08:27.45 &   +51:30:24.01 &   4.802 &     13.04 &   $10014 \pm 340$ &  $4.0 \pm 0.14$ &  $2.2^{+0.14}_{-0.12}$ &   $190^{+127}_{-116}$ &       1 \\
   HIP 39906 &      B5V &    08:09:01.64 &   -19:14:42.05 &   4.390 &      7.01 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.63}_{-0.54}$ &       $11^{+15}_{-5}$ &       2 \\
   HIP 40429 &      A2V &    08:15:15.92 &   -62:54:56.32 &   5.160 &     12.90 &    $8725 \pm 297$ &  $3.9 \pm 0.14$ &  $2.1^{+0.20}_{-0.17}$ &     $635^{+88}_{-88}$ &       1 \\
   HIP 40706 &      A8V &    08:18:33.31 &   -36:39:33.44 &   4.400 &     34.93 &    $7986 \pm 272$ &  $4.3 \pm 0.14$ &  $1.7^{+0.08}_{-0.06}$ &   $540^{+282}_{-330}$ &       1 \\
   HIP 40881 &    B9.5V &    08:20:32.14 &   +24:01:20.32 &   5.930 &      7.13 &   $9500 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.1^{+0.39}_{-0.32}$ &    $105^{+287}_{-93}$ &       2 \\
   HIP 41039 &      B1V &    08:22:31.69 &   -48:29:25.36 &   4.820 &      1.90 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 41307 &     A0Va &    08:25:39.63 &   -03:54:23.12 &   3.900 &     26.66 &   $10281 \pm 350$ &  $4.2 \pm 0.14$ &  $2.3^{+0.12}_{-0.10}$ &   $201^{+109}_{-120}$ &       1 \\
   HIP 42090 &    A2Vnn &    08:34:43.88 &   +36:25:10.63 &   5.755 &      9.19 &   $9000 \pm 1000$ &  $4.0 \pm 0.25$ &  $1.9^{+0.36}_{-0.31}$ &   $133^{+379}_{-119}$ &       2 \\
   HIP 42129 &      B3V &    08:35:15.56 &   -58:13:29.05 &   5.241 &      3.64 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.61}_{-0.56}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 42313 &    A0Vnn &    08:37:39.37 &   +05:42:13.61 &   4.137 &     20.34 &   $11055 \pm 376$ &  $4.0 \pm 0.14$ &  $2.9^{+0.23}_{-0.19}$ &     $260^{+29}_{-44}$ &       1 \\
   HIP 42334 &      A0V &    08:37:52.15 &   -26:15:18.01 &   5.270 &     14.07 &   $11614 \pm 395$ &  $4.3 \pm 0.14$ &  $2.7^{+0.12}_{-0.10}$ &      $79^{+71}_{-51}$ &       1 \\
   HIP 43142 &      A3V &    08:47:14.99 &   -01:53:49.31 &   5.279 &      8.12 &    $9000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.0^{+0.21}_{-0.18}$ &     $94^{+314}_{-83}$ &       2 \\
   HIP 44127 &      A7V &    08:59:12.45 &   +48:02:30.57 &   3.140 &     68.92 &    $8233 \pm 280$ &  $4.4 \pm 0.14$ &  $1.7^{+0.09}_{-0.07}$ &   $474^{+257}_{-294}$ &       1 \\
   HIP 44307 &      A2V &    09:01:24.13 &   +32:15:08.26 &   5.870 &      6.99 &   $8500 \pm 1000$ &  $4.0 \pm 0.25$ &  $1.7^{+0.33}_{-0.28}$ &   $162^{+537}_{-148}$ &       2 \\
   HIP 45336 &    B9.5V &    09:14:21.86 &   +02:18:51.34 &   3.880 &     28.74 &   $10826 \pm 368$ &  $4.2 \pm 0.14$ &  $2.5^{+0.12}_{-0.10}$ &     $111^{+95}_{-71}$ &       1 \\
   HIP 45344 &      B4V &    09:14:24.48 &   -43:13:38.97 &   5.250 &      5.33 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.62}_{-0.57}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 45688 &      A1V &    09:18:50.64 &   +36:48:09.33 &   3.820 &     26.13 &    $8862 \pm 301$ &  $3.9 \pm 0.14$ &  $1.9^{+0.08}_{-0.07}$ &   $255^{+202}_{-160}$ &       1 \\
   HIP 46225 &      A4V &    09:25:27.23 &   -61:57:01.72 &   5.781 &     12.92 &    $8577 \pm 292$ &  $4.0 \pm 0.14$ &  $1.8^{+0.10}_{-0.07}$ &   $418^{+206}_{-252}$ &       1 \\
   HIP 46283 &      B6V &    09:26:17.96 &   -53:22:44.07 &   5.088 &      7.60 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.54}_{-0.50}$ &       $13^{+20}_{-8}$ &       2 \\
   HIP 46897 &    B9.5V &    09:33:26.05 &   -22:51:49.99 &   5.911 &     12.46 &   $10075 \pm 343$ &  $4.2 \pm 0.14$ &  $2.3^{+0.16}_{-0.13}$ &    $286^{+93}_{-140}$ &       1 \\
   HIP 47006 &     A0Vn &    09:34:49.43 &   +52:03:05.32 &   4.479 &     12.44 &    $9757 \pm 332$ &  $3.9 \pm 0.14$ &  $2.1^{+0.10}_{-0.08}$ &   $181^{+138}_{-114}$ &       1 \\
   HIP 47175 &      A7V &    09:36:49.54 &   -49:21:18.09 &   4.350 &     30.94 &    $8331 \pm 283$ &  $4.3 \pm 0.14$ &  $1.8^{+0.11}_{-0.09}$ &   $453^{+256}_{-278}$ &       1 \\
   HIP 50303 &     A0Vn &    10:16:14.43 &   +29:18:37.81 &   5.490 &     12.51 &   $10377 \pm 353$ &  $4.1 \pm 0.14$ &  $2.4^{+0.13}_{-0.11}$ &   $204^{+103}_{-116}$ &       1 \\
   HIP 50860 &      A6V &    10:23:06.33 &   +33:54:29.31 &   5.900 &     13.46 &    $8327 \pm 283$ &  $4.1 \pm 0.14$ &  $1.8^{+0.11}_{-0.09}$ &   $591^{+183}_{-303}$ &       1 \\
   HIP 51362 &    B9.5V &    10:29:28.70 &   -02:44:20.69 &   5.180 &     10.13 &   $10899 \pm 371$ &  $4.0 \pm 0.14$ &  $2.6^{+0.14}_{-0.12}$ &    $182^{+81}_{-101}$ &       1 \\
   HIP 51685 &     A2Vn &    10:33:30.91 &   +34:59:19.30 &   5.580 &      4.77 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.42}_{-0.38}$ &   $406^{+134}_{-172}$ &       2 \\
   HIP 52422 &     A4Vn &    10:43:01.88 &   +26:19:32.09 &   5.517 &     20.67 &    $8170 \pm 278$ &  $4.3 \pm 0.14$ &  $1.7^{+0.08}_{-0.06}$ &   $432^{+278}_{-278}$ &       1 \\
   HIP 52457 &     A3Vn &    10:43:24.96 &   +23:11:18.25 &   5.075 &     14.23 &    $9902 \pm 337$ &  $4.0 \pm 0.14$ &  $2.2^{+0.12}_{-0.10}$ &   $228^{+126}_{-137}$ &       1 \\
   HIP 52638 &     A1Vn &    10:45:51.89 &   +30:40:56.33 &   5.349 &      8.53 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.21}$ &     $69^{+199}_{-59}$ &       2 \\
   HIP 52678 &    B6Vnn &    10:46:16.56 &   -64:30:52.41 &   5.340 &      6.85 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.4^{+0.55}_{-0.48}$ &       $13^{+20}_{-7}$ &       2 \\
   HIP 52736 &   B2.5Vn &    10:46:51.22 &   -64:23:00.50 &   4.850 &      6.79 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.5^{+0.66}_{-0.62}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 52911 &      A2V &    10:49:15.43 &   +10:32:42.73 &   5.314 &      8.57 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.44}_{-0.37}$ &   $406^{+138}_{-172}$ &       2 \\
     HR 4259 &      A1V &    10:55:36.80 &   +24:44:59.00 &   4.500 &   \nodata &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.44}_{-0.36}$ &   $411^{+137}_{-168}$ &       2 \\
   HIP 54849 &      A0V &    11:13:45.55 &   -00:04:10.20 &   5.399 &      6.18 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.24}_{-0.21}$ &     $72^{+197}_{-62}$ &       2 \\
   HIP 55266 &      A2V &    11:19:07.90 &   +38:11:08.03 &   4.800 &     17.00 &    $8720 \pm 296$ &  $4.5 \pm 0.14$ &  $1.9^{+0.10}_{-0.08}$ &   $346^{+208}_{-217}$ &       1 \\
   HIP 55434 &    B9.5V &  11:21:08.1934 &  +06:01:45.571 &   4.044 &     14.82 &   $10618 \pm 361$ &  $3.9 \pm 0.14$ &  $2.7^{+0.23}_{-0.19}$ &     $304^{+32}_{-38}$ &       1 \\
   HIP 56034 &      A2V &    11:29:04.12 &   +39:20:13.11 &   5.354 &     15.33 &   $10000 \pm 500$ &  $4.5 \pm 0.25$ &  $2.2^{+0.21}_{-0.21}$ &     $37^{+118}_{-29}$ &       2 \\
   HIP 56633 &    B9.5V &    11:36:40.91 &   -09:48:08.09 &   4.682 &     11.63 &   $11524 \pm 392$ &  $4.0 \pm 0.14$ &  $2.8^{+0.15}_{-0.12}$ &     $145^{+66}_{-82}$ &       1 \\
   HIP 57328 &      A4V &    11:45:17.04 &   +08:15:29.21 &   4.845 &     26.73 &    $8298 \pm 282$ &  $4.3 \pm 0.14$ &  $1.9^{+0.17}_{-0.14}$ &   $757^{+105}_{-135}$ &       1 \\
   HIP 58590 &      A5V &    12:00:52.39 &   +06:36:51.56 &   4.659 &      8.49 &    $8000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.0^{+0.41}_{-0.35}$ &   $618^{+265}_{-300}$ &       2 \\
   HIP 59394 &      A1V &    12:11:03.84 &   -23:36:08.72 &   5.470 &     17.00 &    $9671 \pm 329$ &  $4.3 \pm 0.14$ &  $2.1^{+0.10}_{-0.09}$ &   $203^{+144}_{-128}$ &       1 \\
   HIP 59449 &      B3V &    12:11:39.12 &   -52:22:06.44 &   3.960 &      8.61 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.61}_{-0.55}$ &       $11^{+14}_{-5}$ &       2 \\
   HIP 59819 &      A3V &    12:16:00.19 &   +14:53:56.65 &   5.090 &     16.42 &    $9528 \pm 324$ &  $4.0 \pm 0.14$ &  $2.2^{+0.16}_{-0.14}$ &    $394^{+95}_{-165}$ &       1 \\
   HIP 60009 &    B2.5V &    12:18:26.25 &   -64:00:11.05 &   4.050 &      9.12 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.65}_{-0.66}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 60030 &     A5Vn &    12:18:40.32 &   -00:47:13.87 &   5.908 &      8.73 &   $7750 \pm 1000$ &  $3.5 \pm 0.25$ &  $1.9^{+0.50}_{-0.39}$ &   $778^{+807}_{-403}$ &       2 \\
   HIP 60595 &      A1V &    12:25:11.76 &   -11:36:38.12 &   5.949 &     14.18 &   $10396 \pm 353$ &  $4.4 \pm 0.14$ &  $2.4^{+0.12}_{-0.10}$ &   $161^{+111}_{-100}$ &       1 \\
   HIP 60710 &     B3Vn &    12:26:31.76 &   -51:27:02.29 &   4.805 &      7.28 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.5^{+0.67}_{-0.63}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 60957 &      A3V &    12:29:43.24 &   +20:53:45.99 &   5.680 &     11.90 &    $9079 \pm 309$ &  $3.9 \pm 0.14$ &  $2.1^{+0.15}_{-0.12}$ &   $480^{+100}_{-192}$ &       1 \\
   HIP 61558 &      A3V &    12:36:47.35 &   -05:49:54.84 &   5.880 &     14.49 &    $9600 \pm 326$ &  $4.2 \pm 0.14$ &  $2.3^{+0.19}_{-0.16}$ &     $436^{+57}_{-75}$ &       1 \\
   HIP 61622 &      A2V &    12:37:42.16 &   -48:32:28.69 &   3.860 &     24.85 &   $10533 \pm 358$ &  $4.0 \pm 0.14$ &  $2.3^{+0.25}_{-0.23}$ &   $158^{+137}_{-102}$ &       1 \\
   HIP 62541 &      A1V &    12:48:54.21 &   +14:07:21.31 &   5.702 &      8.17 &   $9250 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.0^{+0.36}_{-0.31}$ &   $124^{+326}_{-110}$ &       2 \\
   HIP 62576 &      A2V &    12:49:17.45 &   +27:33:08.57 &   5.780 &     10.76 &    $9955 \pm 338$ &  $4.1 \pm 0.14$ &  $2.2^{+0.12}_{-0.11}$ &   $188^{+132}_{-117}$ &       1 \\
   HIP 63724 &      A0V &    13:03:33.31 &   -49:31:38.15 &   4.830 &     14.79 &   $10462 \pm 356$ &  $4.1 \pm 0.14$ &  $2.4^{+0.12}_{-0.10}$ &    $154^{+106}_{-96}$ &       1 \\
   HIP 63945 &      B5V &    13:06:16.70 &   -48:27:47.85 &   4.694 &      8.36 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.61}_{-0.56}$ &       $11^{+13}_{-6}$ &       2 \\
   HIP 65198 &      A2V &    13:21:41.64 &   +02:05:14.07 &   5.693 &     14.77 &   $10291 \pm 350$ &  $4.2 \pm 0.14$ &  $2.3^{+0.11}_{-0.10}$ &   $174^{+111}_{-108}$ &       1 \\
   HIP 65477 &      A5V &    13:25:13.54 &   +54:59:16.65 &   4.010 &     39.91 &    $8221 \pm 280$ &  $4.2 \pm 0.14$ &  $1.8^{+0.16}_{-0.13}$ &   $754^{+157}_{-222}$ &       1 \\
   HIP 65728 &     A1Vn &    13:28:27.09 &   +59:56:44.83 &   5.400 &     14.01 &   $10388 \pm 353$ &  $4.2 \pm 0.14$ &  $2.7^{+0.24}_{-0.19}$ &     $331^{+39}_{-45}$ &       1 \\
   HIP 66249 &    A2Van &    13:34:41.74 &   -00:35:45.38 &   3.380 &     44.03 &    $8542 \pm 290$ &  $4.1 \pm 0.14$ &  $1.9^{+0.15}_{-0.12}$ &   $631^{+113}_{-226}$ &       1 \\
   HIP 66798 &      A2V &    13:41:29.89 &   +64:49:20.68 &   5.850 &     14.66 &    $9347 \pm 318$ &  $4.2 \pm 0.14$ &  $2.1^{+0.11}_{-0.09}$ &   $291^{+153}_{-170}$ &       1 \\
   HIP 66821 &   B8.5Vn &    13:41:44.77 &   -54:33:33.93 &   5.010 &     12.02 &   $13141 \pm 447$ &  $4.4 \pm 0.14$ &  $3.7^{+0.29}_{-0.24}$ &     $138^{+12}_{-20}$ &       1 \\
   HIP 67143 &      A0V &    13:45:36.89 &   -26:06:57.63 &   5.805 &     11.49 &   $10338 \pm 351$ &  $4.3 \pm 0.14$ &  $2.3^{+0.11}_{-0.09}$ &    $143^{+112}_{-93}$ &       1 \\
   HIP 67194 &      A5V &    13:46:13.55 &   +41:05:19.48 &   5.891 &     19.03 &    $7953 \pm 270$ &  $4.3 \pm 0.14$ &  $1.7^{+0.10}_{-0.09}$ &   $504^{+325}_{-319}$ &       1 \\
   HIP 67782 &     A8IV &    13:53:10.28 &   +28:38:53.28 &   5.911 &     15.21 &    $8051 \pm 274$ &  $4.1 \pm 0.14$ &  $1.8^{+0.16}_{-0.12}$ &   $846^{+121}_{-164}$ &       1 \\
   HIP 68092 &      A8V &    13:56:27.88 &   +01:03:02.09 &   5.906 &     11.08 &    $7866 \pm 267$ &  $3.9 \pm 0.14$ &  $1.6^{+0.07}_{-0.07}$ &   $410^{+331}_{-274}$ &       1 \\
   HIP 68520 &      A3V &    14:01:38.79 &   +01:32:40.31 &   4.244 &     14.50 &    $8413 \pm 286$ &  $3.4 \pm 0.14$ &  $2.0^{+0.18}_{-0.14}$ &   $722^{+118}_{-129}$ &       1 \\
   HIP 70327 &      A0V &    14:23:22.70 &   +08:26:47.84 &   5.120 &     15.17 &    $9678 \pm 329$ &  $4.0 \pm 0.14$ &  $2.2^{+0.12}_{-0.09}$ &   $261^{+133}_{-154}$ &       1 \\
   HIP 70384 &      A3V &    14:24:00.88 &   +08:14:38.30 &   5.935 &      7.38 &    $9000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.0^{+0.22}_{-0.18}$ &     $99^{+316}_{-87}$ &       2 \\
   HIP 70400 &      A3V &    14:24:11.34 &   +05:49:12.47 &   5.103 &     20.51 &    $8490 \pm 289$ &  $4.2 \pm 0.14$ &  $1.8^{+0.08}_{-0.07}$ &   $336^{+238}_{-215}$ &       1 \\
   HIP 70915 &     B8Vn &    14:30:08.63 &   -45:19:16.88 &   5.500 &      6.84 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.42}_{-0.37}$ &      $24^{+65}_{-17}$ &       2 \\
     HR 5476 &      A6V &    14:40:43.94 &   +16:25:03.90 &   5.820 &   \nodata &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 71865 &      B3V &    14:41:57.59 &   -37:47:36.59 &   4.000 &      9.62 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.66}_{-0.64}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 71974 &    B9.5V &    14:43:13.55 &   -24:59:51.91 &   5.730 &      6.56 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.26}_{-0.22}$ &     $65^{+210}_{-55}$ &       2 \\
   HIP 72104 &    A0Vnn &    14:44:59.20 &   -35:11:30.57 &   4.923 &     15.22 &    $9618 \pm 327$ &  $3.8 \pm 0.14$ &  $2.1^{+0.12}_{-0.09}$ &   $239^{+143}_{-144}$ &       1 \\
   HIP 72154 &    B9.5V &    14:45:30.20 &   +00:43:02.19 &   5.673 &      6.61 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.24}_{-0.21}$ &     $75^{+199}_{-65}$ &       2 \\
   HIP 72250 &      A1V &    14:46:28.99 &   -47:26:28.02 &   5.740 &     10.67 &    $9784 \pm 333$ &  $4.1 \pm 0.14$ &  $2.4^{+0.21}_{-0.18}$ &     $411^{+49}_{-59}$ &       1 \\
   HIP 72378 &      B9V &    14:47:57.56 &   -26:38:46.16 &   5.768 &      7.29 &   $9500 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.0^{+0.39}_{-0.31}$ &    $103^{+291}_{-91}$ &       2 \\
   HIP 72552 &      A4V &    14:49:58.40 &   +28:36:57.00 &   5.800 &     10.16 &    $9591 \pm 326$ &  $3.9 \pm 0.14$ &  $2.2^{+0.14}_{-0.11}$ &    $359^{+95}_{-168}$ &       1 \\
   HIP 73049 &      A0V &    14:55:44.71 &   -33:51:20.82 &   5.318 &     12.85 &   $10108 \pm 344$ &  $4.3 \pm 0.14$ &  $2.4^{+0.18}_{-0.14}$ &    $337^{+53}_{-113}$ &       1 \\
     HR 5605 &      B5V &    15:05:07.18 &   -47:03:04.00 &   4.720 &   \nodata &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.53}_{-0.52}$ &       $14^{+21}_{-8}$ &       2 \\
   HIP 74117 &      B3V &    15:08:50.62 &   -45:16:47.49 &   4.070 &      4.20 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.67}_{-0.66}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 74689 &      A4V &    15:15:49.08 &   +00:22:19.70 &   5.629 &     21.67 &    $8255 \pm 281$ &  $4.4 \pm 0.14$ &  $1.9^{+0.17}_{-0.14}$ &   $780^{+102}_{-113}$ &       1 \\
   HIP 75178 &     B9Vn &    15:21:48.58 &   +32:56:01.30 &   5.375 &     12.45 &   $12140 \pm 413$ &  $4.3 \pm 0.14$ &  $3.3^{+0.26}_{-0.21}$ &     $184^{+19}_{-28}$ &       1 \\
   HIP 75304 &      B4V &    15:23:09.35 &   -36:51:30.55 &   4.540 &      6.28 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.58}_{-0.55}$ &       $11^{+13}_{-6}$ &       2 \\
   HIP 76267 &      A0V &    15:34:41.27 &   +26:42:52.89 &   2.240 &     43.46 &   $10342 \pm 352$ &  $4.0 \pm 0.14$ &  $2.5^{+0.19}_{-0.16}$ &    $313^{+53}_{-100}$ &       1 \\
   HIP 76600 &    B2.5V &    15:38:39.37 &   -29:46:39.89 &   3.644 &      8.89 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.63}_{-0.65}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 76852 &      A1V &    15:41:33.05 &   +19:40:13.44 &   4.509 &     17.16 &    $9366 \pm 318$ &  $4.1 \pm 0.14$ &  $2.0^{+0.09}_{-0.08}$ &   $190^{+162}_{-123}$ &       1 \\
   HIP 77233 &      A3V &    15:46:11.25 &   +15:25:18.59 &   3.670 &     21.03 &    $8928 \pm 304$ &  $3.3 \pm 0.14$ &  $1.9^{+0.10}_{-0.08}$ &   $332^{+191}_{-201}$ &       1 \\
   HIP 77336 &      A3V &    15:47:17.32 &   +14:06:55.26 &   5.712 &     13.04 &    $8917 \pm 303$ &  $4.0 \pm 0.14$ &  $2.9^{+0.23}_{-0.21}$ &     $403^{+70}_{-75}$ &       1 \\
   HIP 77516 &      A0V &    15:49:37.21 &   -03:25:48.73 &   3.530 &     19.23 &   $10000 \pm 500$ &  $4.5 \pm 0.25$ &  $2.2^{+0.21}_{-0.19}$ &     $37^{+113}_{-28}$ &       2 \\
   HIP 77635 &   B1.5Vn &    15:50:58.74 &   -25:45:04.66 &   4.638 &      6.59 &  $24000 \pm 1000$ &  $4.5 \pm 0.25$ &  $9.0^{+0.60}_{-0.67}$ &         $7^{+4}_{-2}$ &       2 \\
   HIP 77858 &      B5V &    15:53:53.92 &   -24:31:59.37 &   5.376 &      7.76 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.4^{+0.56}_{-0.49}$ &       $13^{+21}_{-8}$ &       2 \\
   HIP 78105 &      A3V &    15:56:53.50 &   -33:57:58.01 &   5.080 &     23.60 &    $9206 \pm 313$ &  $4.1 \pm 0.14$ &  $2.0^{+0.12}_{-0.10}$ &   $291^{+173}_{-171}$ &       1 \\
   HIP 78106 &      B9V &    15:56:54.12 &   -33:57:51.34 &   5.550 &     21.71 &    $9725 \pm 331$ &  $4.2 \pm 0.14$ &  $2.4^{+0.19}_{-0.15}$ &     $412^{+50}_{-83}$ &       1 \\
   HIP 78265 &      B1V &  15:58:51.1132 &  -26:06:50.788 &   2.910 &      5.57 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 78554 &      A3V &    16:02:17.69 &   +22:48:16.03 &   4.817 &     18.22 &    $9226 \pm 314$ &  $4.1 \pm 0.14$ &  $2.1^{+0.15}_{-0.13}$ &    $464^{+88}_{-171}$ &       1 \\
   HIP 78820 &      B1V &    16:05:26.23 &   -19:48:19.63 &   2.620 &      8.07 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 78821 &      B2V &    16:05:26.57 &   -19:48:06.85 &   4.890 &      8.19 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
     HR 6025 &     A0Vn &  16:06:19.6762 &  +67:48:36.480 &   5.439 &     12.85 &   $9250 \pm 1000$ &  $4.5 \pm 0.25$ &  $1.9^{+0.32}_{-0.31}$ &     $48^{+210}_{-39}$ &       2 \\
   HIP 78918 &   B2.5Vn &    16:06:35.55 &   -36:48:08.26 &   4.205 &      7.87 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.68}_{-0.63}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 78933 &      B1V &  16:06:48.4269 &  -20:40:09.090 &   3.970 &      6.92 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 79005 &      B9V &    16:07:36.42 &   -12:44:43.46 &   5.757 &      8.90 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.46}_{-0.42}$ &    $281^{+84}_{-111}$ &       2 \\
   HIP 79007 &      A7V &    16:07:37.54 &   +09:53:30.27 &   5.639 &     10.10 &    $7893 \pm 268$ &  $3.9 \pm 0.14$ &  $1.6^{+0.08}_{-0.07}$ &   $463^{+324}_{-301}$ &       1 \\
   HIP 79199 &      B8V &    16:09:52.59 &   -33:32:44.90 &   5.496 &      8.00 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.40}_{-0.39}$ &      $25^{+65}_{-17}$ &       2 \\
   HIP 79387 &      A4V &    16:12:07.32 &   -08:32:51.28 &   5.435 &     12.92 &    $8418 \pm 286$ &  $4.0 \pm 0.14$ &  $1.9^{+0.16}_{-0.12}$ &   $694^{+108}_{-181}$ &       1 \\
   HIP 79404 &      B2V &    16:12:18.20 &   -27:55:34.95 &   4.567 &      6.81 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.71}_{-0.68}$ &         $8^{+6}_{-3}$ &       2 \\
   HIP 79653 &      B8V &    16:15:15.32 &   -47:22:19.27 &   5.124 &      8.46 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.47}_{-0.44}$ &      $17^{+35}_{-11}$ &       2 \\
   HIP 80460 &      A5V &    16:25:24.17 &   +37:23:38.68 &   5.540 &     13.43 &    $7810 \pm 266$ &  $3.7 \pm 0.14$ &  $1.7^{+0.11}_{-0.08}$ &   $783^{+229}_{-412}$ &       1 \\
   HIP 80815 &      B3V &    16:30:12.48 &   -25:06:54.80 &   4.790 &      7.89 &  $22000 \pm 1000$ &  $4.5 \pm 0.25$ &  $7.8^{+0.76}_{-0.71}$ &         $8^{+6}_{-3}$ &       2 \\
   HIP 80883 &      A0V &    16:30:54.82 &   +01:59:02.12 &   3.900 &     18.84 &    $9000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.0^{+0.22}_{-0.18}$ &     $97^{+294}_{-86}$ &       2 \\
   HIP 80991 &      A2V &    16:32:25.68 &   +60:49:23.96 &   5.910 &      8.79 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.42}_{-0.39}$ &   $405^{+134}_{-176}$ &       2 \\
   HIP 81126 &      B9V &    16:34:06.18 &   +42:26:13.34 &   4.196 &     10.36 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 81641 &      A1V &    16:40:38.69 &   +04:13:11.23 &   5.772 &     11.11 &   $10572 \pm 359$ &  $4.1 \pm 0.14$ &  $2.6^{+0.21}_{-0.17}$ &     $301^{+38}_{-66}$ &       1 \\
   HIP 82350 &      A1V &    16:49:34.66 &   +13:15:40.10 &   5.910 &      9.99 &    $9630 \pm 327$ &  $4.1 \pm 0.14$ &  $2.1^{+0.10}_{-0.08}$ &   $207^{+145}_{-130}$ &       1 \\
   HIP 82514 &    B1.5V &  16:51:52.2311 &  -38:02:50.569 &   2.980 &      6.51 &  $26000 \pm 1000$ &  $4.5 \pm 0.25$ &  $9.6^{+0.25}_{-0.43}$ &         $5^{+3}_{-1}$ &       2 \\
   HIP 82673 &      B8V &    16:54:00.47 &   +10:09:55.30 &   4.380 &     13.30 &  $12000 \pm 1000$ &  $3.8 \pm 0.25$ &  $3.1^{+0.58}_{-0.45}$ &     $118^{+82}_{-98}$ &       2 \\
   HIP 83635 &      B1V &    17:05:32.26 &   -00:53:31.45 &   5.610 &      2.92 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 84606 &      A2V &    17:17:40.25 &   +37:17:29.40 &   4.650 &     18.59 &    $9702 \pm 330$ &  $4.0 \pm 0.14$ &  $2.1^{+0.10}_{-0.09}$ &   $155^{+137}_{-100}$ &       1 \\
   HIP 85290 &     A1Vn &    17:25:41.35 &   +60:02:54.23 &   5.645 &     10.32 &   $10989 \pm 374$ &  $4.0 \pm 0.14$ &  $2.5^{+0.12}_{-0.11}$ &     $110^{+88}_{-70}$ &       1 \\
   HIP 85379 &      A4V &    17:26:44.24 &   +48:15:36.23 &   5.830 &      8.68 &    $8000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.0^{+0.42}_{-0.32}$ &   $621^{+266}_{-320}$ &       2 \\
   HIP 85385 &      B5V &    17:26:49.13 &   +20:04:51.52 &   5.510 &      5.59 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
   HIP 85537 &     A8Vp &    17:28:49.65 &   +00:19:50.25 &   5.424 &     16.77 &    $7624 \pm 259$ &  $3.8 \pm 0.14$ &  $1.8^{+0.16}_{-0.14}$ &  $1064^{+148}_{-159}$ &       1 \\
   HIP 85727 &     B8Vn &    17:31:05.91 &   -60:41:01.85 &   3.620 &     16.48 &   $12740 \pm 433$ &  $4.0 \pm 0.14$ &  $3.2^{+0.15}_{-0.13}$ &      $92^{+48}_{-54}$ &       1 \\
   HIP 85922 &      A5V &    17:33:29.85 &   -05:44:41.29 &   5.619 &     20.79 &    $8000 \pm 500$ &  $4.0 \pm 0.25$ &  $1.7^{+0.19}_{-0.16}$ &   $117^{+539}_{-105}$ &       2 \\
   HIP 85998 &      A2V &    17:34:36.69 &   +09:35:12.12 &   5.810 &      9.66 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.44}_{-0.37}$ &   $399^{+137}_{-200}$ &       2 \\
   HIP 86019 &     B8Vn &    17:34:46.35 &   -11:14:31.19 &   5.537 &      8.11 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.42}_{-0.37}$ &      $25^{+62}_{-17}$ &       2 \\
   HIP 86782 &      A2V &    17:43:59.18 &   +53:48:06.17 &   5.760 &      6.12 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.22}$ &     $58^{+207}_{-49}$ &       2 \\
   HIP 87108 &     A0Vn &    17:47:53.56 &   +02:42:26.20 &   3.750 &     31.73 &   $10075 \pm 343$ &  $4.2 \pm 0.14$ &  $2.3^{+0.15}_{-0.12}$ &   $258^{+101}_{-138}$ &       1 \\
   HIP 88116 &       A0 &    17:59:47.56 &   -23:48:58.09 &   4.731 &      7.85 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.20}$ &     $76^{+200}_{-66}$ &       2 \\
   HIP 88290 &     A2Vn &    18:01:45.20 &   +01:18:18.28 &   4.439 &     11.15 &   $10010 \pm 340$ &  $3.8 \pm 0.14$ &  $2.3^{+0.12}_{-0.11}$ &   $243^{+112}_{-140}$ &       1 \\
   HIP 88818 &      A3V &    18:07:49.45 &   +26:06:04.13 &   5.870 &     20.16 &    $9162 \pm 312$ &  $4.3 \pm 0.14$ &  $2.7^{+0.24}_{-0.23}$ &     $442^{+86}_{-70}$ &       1 \\
   HIP 88817 &      A3V &    18:07:49.50 &   +26:05:50.40 &   5.900 &     25.92 &    $9162 \pm 312$ &  $4.3 \pm 0.14$ &  $2.1^{+0.15}_{-0.12}$ &    $481^{+83}_{-173}$ &       1 \\
   HIP 89156 &      A3V &    18:11:45.12 &   +33:26:49.41 &   5.970 &      4.21 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.42}_{-0.39}$ &   $410^{+135}_{-157}$ &       2 \\
   HIP 89935 &      A7V &    18:21:01.02 &   +28:52:11.83 &   5.129 &     12.40 &    $7594 \pm 258$ &  $3.4 \pm 0.14$ &  $2.3^{+0.10}_{-0.13}$ &    $727^{+108}_{-70}$ &       1 \\
   HIP 90052 &      A2V &    18:22:35.32 &   +12:01:46.85 &   5.980 &      7.42 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.43}_{-0.40}$ &   $404^{+138}_{-175}$ &       2 \\
   HIP 90762 &      A2V &    18:31:04.45 &   +16:55:42.80 &   5.760 &      7.49 &   $8750 \pm 1000$ &  $3.5 \pm 0.25$ &  $2.2^{+0.54}_{-0.46}$ &   $482^{+375}_{-247}$ &       2 \\
   HIP 90887 &     A3Vn &    18:32:21.33 &   -39:42:14.40 &   5.162 &     14.23 &    $9403 \pm 320$ &  $4.1 \pm 0.14$ &  $2.1^{+0.14}_{-0.12}$ &   $335^{+143}_{-187}$ &       1 \\
   HIP 91118 &     A0Vn &    18:35:12.60 &   +18:12:12.28 &   5.790 &      5.04 &   $9500 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.1^{+0.39}_{-0.32}$ &    $111^{+276}_{-98}$ &       2 \\
   HIP 91875 &     A2Vn &    18:43:46.94 &   -38:19:24.39 &   5.120 &     15.89 &    $9287 \pm 316$ &  $4.0 \pm 0.14$ &  $2.1^{+0.12}_{-0.10}$ &   $339^{+145}_{-189}$ &       1 \\
   HIP 92027 &      A1V &    18:45:28.36 &   +05:30:00.44 &   5.830 &      5.30 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.43}_{-0.37}$ &   $406^{+140}_{-167}$ &       2 \\
   HIP 92312 &      A1V &    18:48:53.39 &   +19:19:43.40 &   5.894 &     11.41 &    $9923 \pm 337$ &  $4.2 \pm 0.14$ &  $2.2^{+0.10}_{-0.09}$ &   $164^{+128}_{-105}$ &       1 \\
   HIP 92728 &    B2.5V &    18:53:43.56 &   +36:58:18.19 &   5.569 &      3.29 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.68}_{-0.61}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 92855 &      B2V &    18:55:15.93 &   -26:17:48.21 &   2.058 &     14.32 &   $19192 \pm 653$ &  $4.3 \pm 0.14$ &  $6.3^{+0.38}_{-0.28}$ &       $29^{+6}_{-12}$ &       1 \\
   HIP 92946 &      A5V &    18:56:13.18 &   +04:12:12.91 &   4.620 &     21.09 &    $7968 \pm 271$ &  $4.1 \pm 0.14$ &  $1.6^{+0.19}_{-0.23}$ &   $658^{+683}_{-409}$ &       1 \\
   HIP 93225 &      B4V &    18:59:23.80 &   -12:50:25.86 &   5.516 &      6.84 &  $15000 \pm 1000$ &  $3.8 \pm 0.25$ &  $4.4^{+0.71}_{-0.55}$ &      $54^{+34}_{-42}$ &       2 \\
   HIP 93393 &      B5V &    19:01:17.36 &   +26:17:29.08 &   5.683 &      5.02 &  $18000 \pm 1000$ &  $4.5 \pm 0.25$ &  $5.4^{+0.60}_{-0.55}$ &       $11^{+13}_{-5}$ &       2 \\
   HIP 93580 &      A4V &    19:03:32.25 &   +01:49:07.57 &   5.826 &     18.22 &    $8224 \pm 280$ &  $4.3 \pm 0.14$ &  $1.8^{+0.12}_{-0.10}$ &   $590^{+236}_{-315}$ &       1 \\
   HIP 93713 &     A0Vn &    19:04:55.17 &   +53:23:47.96 &   5.380 &      9.00 &   $9000 \pm 1000$ &  $4.0 \pm 0.25$ &  $1.9^{+0.36}_{-0.31}$ &   $129^{+363}_{-115}$ &       2 \\
   HIP 93747 &     A0Vn &    19:05:24.61 &   +13:51:48.52 &   2.990 &     39.28 &   $11409 \pm 388$ &  $4.2 \pm 0.14$ &  $3.3^{+0.29}_{-0.25}$ &     $231^{+23}_{-26}$ &       1 \\
   HIP 93805 &     B9Vn &    19:06:14.94 &   -04:52:57.20 &   3.430 &     26.37 &   $11962 \pm 407$ &  $4.2 \pm 0.14$ &  $3.5^{+0.33}_{-0.27}$ &     $194^{+20}_{-22}$ &       1 \\
   HIP 94620 &      A4V &    19:15:17.36 &   +21:13:55.62 &   5.654 &     10.89 &    $9190 \pm 312$ &  $4.0 \pm 0.14$ &  $2.0^{+0.09}_{-0.07}$ &   $202^{+169}_{-130}$ &       1 \\
   HIP 94720 &    B9.5V &    19:16:26.79 &   +14:32:40.62 &   5.630 &      7.96 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.20}$ &     $69^{+194}_{-58}$ &       2 \\
   HIP 95241 &      B9V &    19:22:38.30 &   -44:27:32.25 &   4.010 &     10.40 &  $12000 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.9^{+0.50}_{-0.41}$ &     $53^{+108}_{-43}$ &       2 \\
   HIP 95560 &      A0V &    19:26:13.25 &   +20:05:51.84 &   5.594 &     13.72 &   $9250 \pm 1000$ &  $4.5 \pm 0.25$ &  $1.9^{+0.32}_{-0.29}$ &     $45^{+197}_{-37}$ &       2 \\
   HIP 95619 &    B8.5V &    19:26:56.48 &   -29:44:35.62 &   5.650 &     14.30 &   $11997 \pm 408$ &  $4.3 \pm 0.14$ &  $2.9^{+0.15}_{-0.12}$ &     $107^{+60}_{-63}$ &       1 \\
   HIP 95853 &      A5V &    19:29:42.36 &   +51:43:47.21 &   3.769 &     26.88 &    $8216 \pm 279$ &  $3.9 \pm 0.14$ &  $1.8^{+0.14}_{-0.11}$ &   $693^{+186}_{-300}$ &       1 \\
   HIP 96288 &      A2V &    19:34:41.26 &   +42:24:45.04 &   5.348 &      5.66 &    $9000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.4^{+0.43}_{-0.39}$ &   $405^{+137}_{-166}$ &       2 \\
   HIP 96840 &      B5V &    19:41:05.53 &   +13:48:56.45 &   5.988 &      0.49 &  $16000 \pm 1000$ &  $4.0 \pm 0.25$ &  $4.7^{+0.62}_{-0.56}$ &      $26^{+31}_{-18}$ &       2 \\
   HIP 97376 &     B8Vn &    19:47:27.78 &   +38:24:27.41 &   5.826 &      6.90 &  $11500 \pm 2000$ &  $4.5 \pm 0.25$ &  $2.2^{+0.73}_{-0.70}$ &     $34^{+157}_{-26}$ &       2 \\
   HIP 97496 &      A3V &    19:48:58.66 &   +19:08:31.35 &   5.000 &     12.79 &    $8422 \pm 286$ &  $3.9 \pm 0.14$ &  $1.8^{+0.09}_{-0.07}$ &   $435^{+227}_{-265}$ &       1 \\
   HIP 97870 &      B5V &    19:53:17.38 &   +57:31:24.53 &   5.132 &      5.15 &  $16000 \pm 1000$ &  $4.0 \pm 0.25$ &  $4.7^{+0.64}_{-0.53}$ &      $26^{+32}_{-19}$ &       2 \\
   HIP 97966 &     B7Vn &    19:54:37.65 &   -08:13:38.24 &   5.710 &      6.76 &  $14000 \pm 1000$ &  $4.2 \pm 0.25$ &  $3.6^{+0.47}_{-0.46}$ &      $24^{+45}_{-16}$ &       2 \\
   HIP 98055 &     A4Vn &    19:55:37.79 &   +52:26:20.21 &   4.920 &     11.59 &    $9000 \pm 500$ &  $4.5 \pm 0.25$ &  $1.9^{+0.18}_{-0.17}$ &     $47^{+186}_{-38}$ &       2 \\
   HIP 98194 &      B5V &    19:57:13.87 &   +40:22:04.17 &   5.453 &      2.94 &  $16000 \pm 1000$ &  $4.0 \pm 0.25$ &  $4.7^{+0.59}_{-0.54}$ &      $27^{+31}_{-19}$ &       2 \\
   HIP 98325 &     B9Vn &    19:58:37.98 &   +30:59:01.19 &   5.506 &      7.01 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.41}_{-0.37}$ &      $25^{+65}_{-17}$ &       2 \\
   HIP 99080 &      B3V &    20:06:53.41 &   +23:36:51.93 &   5.064 &      6.47 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.68}_{-0.64}$ &         $9^{+9}_{-4}$ &       2 \\
   HIP 99742 &     A1Va &    20:14:16.62 &   +15:11:51.39 &   4.947 &     21.75 &    $9645 \pm 328$ &  $4.3 \pm 0.14$ &  $2.3^{+0.17}_{-0.14}$ &    $412^{+59}_{-116}$ &       1 \\
  HIP 100069 &      B0V &    20:18:06.99 &   +40:43:55.50 &   5.840 &      1.68 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
  HIP 100221 &      B9V &    20:19:36.72 &   +62:15:26.90 &   5.712 &      8.63 &   $10000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.3^{+0.25}_{-0.21}$ &     $70^{+205}_{-60}$ &       2 \\
  HIP 100881 &      B4V &    20:27:19.21 &   -18:12:42.20 &   5.250 &      5.98 &  $14000 \pm 1000$ &  $4.5 \pm 0.25$ &  $3.6^{+0.48}_{-0.45}$ &      $17^{+34}_{-10}$ &       2 \\
  HIP 100907 &      A3V &    20:27:34.26 &   +38:26:25.19 &   5.620 &     12.28 &    $9042 \pm 307$ &  $4.0 \pm 0.14$ &  $2.1^{+0.16}_{-0.13}$ &    $514^{+99}_{-166}$ &       1 \\
  HIP 101123 &      A1V &    20:29:53.91 &   -18:34:59.48 &   5.906 &     15.07 &   $10492 \pm 357$ &  $4.2 \pm 0.14$ &  $2.4^{+0.11}_{-0.09}$ &    $143^{+109}_{-91}$ &       1 \\
  HIP 101589 &      A3V &    20:35:18.54 &   +14:40:27.17 &   4.664 &     14.82 &    $8639 \pm 294$ &  $3.8 \pm 0.14$ &  $2.0^{+0.15}_{-0.12}$ &    $624^{+98}_{-176}$ &       1 \\
  HIP 101716 &      B8V &    20:37:04.67 &   +26:27:43.01 &   5.589 &     10.27 &   $12949 \pm 440$ &  $4.2 \pm 0.14$ &  $3.3^{+0.17}_{-0.14}$ &      $91^{+44}_{-53}$ &       1 \\
  HIP 101867 &      A0V &    20:38:31.34 &   +21:12:04.25 &   4.820 &     14.01 &   $10507 \pm 357$ &  $4.2 \pm 0.14$ &  $3.4^{+0.24}_{-0.30}$ &     $252^{+45}_{-33}$ &       1 \\
  HIP 101909 &      B3V &    20:39:04.97 &   +15:50:17.52 &   5.980 &      2.42 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
  HIP 102487 &      A1V &    20:46:09.99 &   -21:30:50.52 &   5.913 &     12.37 &    $9702 \pm 330$ &  $4.1 \pm 0.14$ &  $2.1^{+0.11}_{-0.09}$ &   $180^{+143}_{-116}$ &       1 \\
  HIP 103298 &      A5V &    20:55:38.57 &   +12:34:06.81 &   5.540 &     16.53 &    $9093 \pm 309$ &  $4.1 \pm 0.14$ &  $2.0^{+0.13}_{-0.11}$ &   $400^{+150}_{-215}$ &       1 \\
  HIP 104019 &      A5V &    21:04:24.30 &   -19:51:17.97 &   4.840 &     20.20 &    $8319 \pm 283$ &  $4.2 \pm 0.14$ &  $1.8^{+0.09}_{-0.07}$ &   $452^{+245}_{-273}$ &       1 \\
  HIP 104105 &     B8Vn &    21:05:29.27 &   +78:07:35.02 &   5.915 &      8.10 &   $9750 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.1^{+0.39}_{-0.32}$ &     $92^{+255}_{-80}$ &       2 \\
  HIP 104139 &      A1V &    21:05:56.83 &   -17:13:58.30 &   4.070 &     20.11 &   $10001 \pm 340$ &  $4.2 \pm 0.14$ &  $2.2^{+0.11}_{-0.09}$ &   $187^{+130}_{-116}$ &       1 \\
  HIP 104365 &      A0V &    21:08:33.63 &   -21:11:37.22 &   5.280 &     18.14 &   $10878 \pm 370$ &  $4.3 \pm 0.14$ &  $2.8^{+0.21}_{-0.17}$ &     $272^{+33}_{-53}$ &       1 \\
  HIP 105140 &      A1V &    21:17:56.28 &   -32:10:21.15 &   4.721 &     17.90 &    $9126 \pm 310$ &  $4.4 \pm 0.14$ &  $2.3^{+0.21}_{-0.18}$ &     $538^{+69}_{-70}$ &       1 \\
  HIP 105282 &      B6V &    21:19:28.75 &   +49:30:37.06 &   5.740 &      5.98 &  $16000 \pm 1000$ &  $4.5 \pm 0.25$ &  $4.5^{+0.53}_{-0.51}$ &       $13^{+21}_{-7}$ &       2 \\
  HIP 105891 &      B6V &    21:26:44.97 &   +52:53:54.73 &   5.989 &      7.28 &  $10500 \pm 2000$ &  $4.0 \pm 0.25$ &  $2.1^{+0.71}_{-0.58}$ &    $102^{+404}_{-89}$ &       2 \\
  HIP 105942 &      B3V &    21:27:21.37 &   +37:07:00.47 &   5.289 &      2.50 &  $20000 \pm 1000$ &  $4.5 \pm 0.25$ &  $6.6^{+0.69}_{-0.64}$ &         $9^{+9}_{-4}$ &       2 \\
  HIP 105966 &      A1V &    21:27:40.06 &   +27:36:30.94 &   5.389 &     17.14 &    $9622 \pm 327$ &  $4.3 \pm 0.14$ &  $2.1^{+0.12}_{-0.10}$ &   $263^{+138}_{-156}$ &       1 \\
  HIP 105972 &      B7V &    21:27:46.14 &   +66:48:32.74 &   5.407 &      4.05 &  $14000 \pm 1000$ &  $4.0 \pm 0.25$ &  $3.8^{+0.53}_{-0.48}$ &      $38^{+54}_{-29}$ &       2 \\
  HIP 106711 &     A3Vn &    21:36:56.98 &   +40:24:48.67 &   5.052 &     15.19 &    $7859 \pm 267$ &  $3.9 \pm 0.14$ &  $1.7^{+0.11}_{-0.08}$ &   $767^{+226}_{-387}$ &       1 \\
  HIP 106786 &      A7V &    21:37:45.11 &   -07:51:15.13 &   4.690 &     18.26 &    $8140 \pm 277$ &  $4.0 \pm 0.14$ &  $1.7^{+0.10}_{-0.07}$ &   $567^{+239}_{-326}$ &       1 \\
  HIP 107517 &      A1V &    21:46:32.10 &   -11:21:57.44 &   5.570 &     11.58 &   $10674 \pm 363$ &  $4.2 \pm 0.14$ &  $2.5^{+0.14}_{-0.11}$ &    $191^{+89}_{-108}$ &       1 \\
  HIP 107608 &      A2V &    21:47:44.15 &   -30:53:53.90 &   5.016 &     10.16 &    $9716 \pm 330$ &  $4.0 \pm 0.14$ &  $2.2^{+0.15}_{-0.12}$ &    $342^{+93}_{-161}$ &       1 \\
  HIP 108294 &     A2Vn &    21:56:22.77 &   -37:15:13.16 &   5.457 &      9.71 &    $9000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.0^{+0.22}_{-0.18}$ &     $93^{+318}_{-81}$ &       2 \\
  HIP 108339 &    A2Vnn &    21:56:56.37 &   +12:04:35.37 &   5.544 &      8.14 &   $9250 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.0^{+0.35}_{-0.31}$ &   $121^{+316}_{-109}$ &       2 \\
  HIP 109056 &     B9Vn &    22:05:34.67 &   +28:57:50.32 &   5.700 &      9.40 &   $9800 \pm 1200$ &  $4.4 \pm 0.25$ &  $2.0^{+0.41}_{-0.36}$ &     $51^{+200}_{-42}$ &       2 \\
  HIP 109139 &      B8V &    22:06:26.23 &   -13:52:10.86 &   4.270 &     18.62 &   $12641 \pm 430$ &  $4.3 \pm 0.14$ &  $3.2^{+0.20}_{-0.16}$ &     $124^{+37}_{-62}$ &       1 \\
  HIP 109521 &      A5V &    22:11:09.89 &   +50:49:24.26 &   5.386 &     17.81 &    $8420 \pm 286$ &  $4.2 \pm 0.14$ &  $2.0^{+0.19}_{-0.16}$ &   $728^{+107}_{-106}$ &       1 \\
  HIP 109831 &    A2Vnn &    22:14:44.37 &   +42:57:14.07 &   5.720 &     12.14 &   $10932 \pm 372$ &  $4.2 \pm 0.14$ &  $2.5^{+0.13}_{-0.12}$ &     $123^{+92}_{-79}$ &       1 \\
  HIP 110838 &     B9Vn &    22:27:19.97 &   -64:57:58.88 &   4.495 &     13.00 &   $11271 \pm 383$ &  $4.1 \pm 0.14$ &  $3.0^{+0.24}_{-0.20}$ &     $243^{+24}_{-33}$ &       1 \\
  HIP 110935 &      A4V &    22:28:37.67 &   -67:29:20.62 &   5.570 &     23.19 &    $8026 \pm 273$ &  $4.3 \pm 0.14$ &  $1.7^{+0.08}_{-0.07}$ &   $447^{+301}_{-291}$ &       1 \\
  HIP 111056 &      A3V &    22:29:52.98 &   +78:49:27.43 &   5.460 &     13.31 &   $8750 \pm 1000$ &  $4.2 \pm 0.25$ &  $1.8^{+0.33}_{-0.30}$ &     $85^{+347}_{-74}$ &       2 \\
  HIP 111068 &    B9.5V &    22:30:01.81 &   +32:34:21.50 &   5.650 &      9.05 &   $10000 \pm 500$ &  $3.5 \pm 0.25$ &  $2.8^{+0.47}_{-0.42}$ &    $280^{+85}_{-112}$ &       2 \\
  HIP 111169 &      A1V &    22:31:17.50 &   +50:16:56.97 &   3.770 &     31.79 &   $10071 \pm 342$ &  $4.3 \pm 0.14$ &  $2.3^{+0.12}_{-0.10}$ &   $202^{+123}_{-123}$ &       1 \\
  HIP 111841 &      O9V &    22:39:15.68 &   +39:03:00.97 &   4.880 &      1.89 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata \\
  HIP 112029 &      B8V &    22:41:27.73 &   +10:49:52.64 &   3.410 &   \nodata &   $11913 \pm 405$ &  $3.8 \pm 0.14$ &  $2.9^{+0.17}_{-0.14}$ &     $135^{+57}_{-76}$ &       1 \\
  HIP 113788 &      A1V &    23:02:36.38 &   +42:45:28.06 & \nodata &      7.74 &    $9000 \pm 500$ &  $4.0 \pm 0.25$ &  $2.0^{+0.22}_{-0.18}$ &    $100^{+309}_{-88}$ &       2 \\
  HIP 114520 &     A5Vn &    23:11:44.19 &   +08:43:12.40 &   5.157 &     13.46 &    $8220 \pm 279$ &  $3.8 \pm 0.14$ &  $1.7^{+0.10}_{-0.08}$ &   $437^{+277}_{-270}$ &       1 \\
  HIP 114822 &      A2V &    23:15:34.26 &   -03:29:46.96 &   5.550 &     14.04 &    $9678 \pm 329$ &  $4.1 \pm 0.14$ &  $2.2^{+0.12}_{-0.10}$ &   $255^{+136}_{-153}$ &       1 \\
  HIP 115115 &      A0V &    23:18:57.68 &   -09:36:38.70 &   5.001 &     12.47 &   $10671 \pm 363$ &  $4.2 \pm 0.14$ &  $2.6^{+0.20}_{-0.16}$ &     $284^{+38}_{-79}$ &       1 \\
  HIP 116247 &      A0V &    23:33:16.62 &   -20:54:52.21 &   4.709 &     11.11 &   $10267 \pm 349$ &  $3.9 \pm 0.14$ &  $2.5^{+0.20}_{-0.16}$ &     $334^{+46}_{-78}$ &       1 \\
  HIP 116582 &      B8V &    23:37:32.04 &   +44:25:44.37 &   5.817 &      4.04 &  $12000 \pm 1000$ &  $4.0 \pm 0.25$ &  $2.9^{+0.46}_{-0.43}$ &     $54^{+104}_{-44}$ &       2 \\
  HIP 116611 &     A0Vn &    23:37:56.80 &   +18:24:02.40 &   5.482 &     14.01 &   $11060 \pm 376$ &  $4.1 \pm 0.14$ &  $2.7^{+0.19}_{-0.15}$ &     $228^{+46}_{-92}$ &       1 \\
  HIP 116631 &      B8V &    23:38:08.20 &   +43:16:05.06 &   4.290 &      6.53 &  $12000 \pm 1000$ &  $3.8 \pm 0.25$ &  $3.1^{+0.55}_{-0.48}$ &     $116^{+82}_{-96}$ &       2 \\
  HIP 116805 &      A0V &    23:40:24.51 &   +44:20:02.16 &   4.140 &     19.37 &   $11903 \pm 405$ &  $4.3 \pm 0.14$ &  $2.9^{+0.17}_{-0.14}$ &     $145^{+51}_{-74}$ &       1 \\
  HIP 116971 &      B9V &    23:42:43.34 &   -14:32:41.65 &   4.483 &     21.96 &   $11108 \pm 378$ &  $4.3 \pm 0.14$ &  $2.6^{+0.15}_{-0.14}$ &     $109^{+90}_{-70}$ &       1 \\
  HIP 117089 &      B9V &    23:44:12.08 &   -18:16:36.97 &   5.235 &      8.61 &  $12000 \pm 1000$ &  $4.5 \pm 0.25$ &  $2.8^{+0.40}_{-0.37}$ &      $24^{+63}_{-17}$ &       2 \\
  HIP 117371 &     A1Vn &    23:47:54.77 &   +67:48:24.51 &   5.048 &     10.76 &   $10308 \pm 350$ &  $3.9 \pm 0.14$ &  $2.5^{+0.19}_{-0.17}$ &     $323^{+53}_{-92}$ &       1 \\
  HIP 117452 &     A0Vn &    23:48:55.55 &   -28:07:48.97 &   4.570 &     23.73 &   $11417 \pm 388$ &  $4.3 \pm 0.14$ &  $3.0^{+0.26}_{-0.22}$ &     $233^{+32}_{-37}$ &       1 \\
  HIP 118121 &      A1V &    23:57:35.08 &   -64:17:53.63 &   5.002 &     21.08 &    $9979 \pm 339$ &  $4.3 \pm 0.14$ &  $2.3^{+0.16}_{-0.13}$ &    $335^{+70}_{-134}$ &       1 \\
  HIP 118243 &      B1V &    23:59:00.54 &   +55:45:17.74 &   4.997 &      0.72 &           \nodata &         \nodata &                \nodata &               \nodata & \nodata

\enddata
\tablecomments{The spectral types, coordinates, V-band magnitudes, and parallax measurements are taken from the Simbad database; the spectral type given is that of the brightest star if part of a known multiple system. The "Ref" column denotes the reference for the stellar effective temperature, surface gravity, mass, and age. The references are: [1]: \citep{David2015}; [2]: This study. We only estimate stellar parameters for the stars in which we detect a companion.}
\label{tab:sample}
\end{deluxetable*}

\clearpage
%\newpage

%\LongTables
\begin{deluxetable}{|l|rlrr|}
\tabletypesize{\small}
\tablewidth{0pt}
\tablecaption{Spectroscopic Observation Log}
\tablehead{
      \colhead{} &  \colhead{MJD} & \colhead{}  & \colhead{exptime} & \colhead{} \\
       \colhead{star} &  \colhead{(JD-2450000)} &   \colhead{instrument} & \colhead{(sec)} & \colhead{snr}
       }

\startdata
     HIP 813 &  6618.58 &       TS23 &      557 &  1810 \\
    HIP 1191 &  6552.60 &     CHIRON &     1200 &   320 \\
    HIP 1366 &  6585.73 &       TS23 &      274 &  1790 \\
    HIP 1647 &  6510.74 &     CHIRON &     1105 &   630 \\
    HIP 2381 &  6874.71 &     CHIRON &      505 &   580 \\
    HIP 2505 &  6586.71 &       TS23 &      235 &  1700 \\
    HIP 2505 &  6962.71 &       TS23 &      804 &  1960 \\
    HIP 2505 &  6945.71 &     IGRINS &      224 &   210 \\
    HIP 2548 &  7240.93 &     IGRINS &      360 &   100 \\
    HIP 2548 &  6516.84 &        HRS &      412 &  1070 \\
    HIP 2912 &  7238.96 &     IGRINS &      367 &   130 \\
    HIP 2912 &  6583.78 &       TS23 &      199 &  2460 \\
    HIP 3300 &  6299.55 &       TS23 &      689 &  1010 \\
    HIP 3478 &  6963.63 &       TS23 &     1200 &  1020 \\
    HIP 3478 &  7288.82 &     IGRINS &      540 &    30 \\
    HIP 3478 &  6669.53 &       TS23 &      743 &  1870 \\
    HIP 3478 &  6945.72 &     IGRINS &      224 &   150 \\
    HIP 3741 &  7174.89 &     CHIRON &     3981 &   430 \\
    HIP 3741 &  7175.90 &     CHIRON &     3561 &   420 \\
    HIP 3881 &  6585.70 &       TS23 &      222 &  1690 \\
    HIP 3881 &  6964.60 &       TS23 &     1200 &   710 \\
    HIP 3881 &  6945.74 &     IGRINS &      224 &   200 \\
    HIP 4436 &  6585.76 &       TS23 &      129 &  1940 \\
    HIP 5131 &  6587.70 &       TS23 &      655 &  1780 \\
    HIP 5131 &  7083.57 &     IGRINS &      930 &    70 \\
    HIP 5131 &  6847.89 &     IGRINS &     1600 &   110 \\
    HIP 5132 &  6587.77 &       TS23 &      535 &  1550 \\
    HIP 5132 &  6847.89 &     IGRINS &      480 &   110 \\
    HIP 5310 &  6945.75 &     IGRINS &      224 &   200 \\
    HIP 5348 &  6325.56 &     CHIRON &      186 &   390 \\
    HIP 5361 &  6523.84 &        HRS &     1000 &   810 \\
    HIP 5518 &  7238.94 &     IGRINS &      540 &    80 \\
    HIP 5518 &  6945.77 &     IGRINS &      224 &   160 \\
    HIP 5626 &  6945.88 &     IGRINS &      224 &   160 \\
    HIP 7345 &  6888.76 &     CHIRON &     1080 &   480 \\
    HIP 7345 &  6887.76 &     CHIRON &      706 &   540 \\
    HIP 8016 &  6522.96 &        HRS &      328 &   580 \\
    HIP 8704 &  6506.90 &        HRS &      365 &   420 \\
    HIP 8704 &  7084.57 &     IGRINS &      960 &    30 \\
    HIP 8704 &  7237.93 &     IGRINS &      720 &    90 \\
      HR 545 &  6668.56 &       TS23 &      597 &  1920 \\
      HR 545 &  6586.75 &       TS23 &      265 &  1710 \\
      HR 545 &  6945.81 &     IGRINS &      224 &   230 \\
    HIP 9312 &  6586.78 &       TS23 &      419 &  1750 \\
    HIP 9564 &  6945.78 &     IGRINS &      224 &   140 \\
      HR 604 &  6585.86 &       TS23 &      349 &  1650 \\
   HIP 10320 &  6532.80 &     CHIRON &     1020 &   550 \\
   HIP 10670 &  6618.68 &       TS23 &      141 &  1810 \\
   HIP 10732 &  6945.82 &     IGRINS &      224 &   180 \\
   HIP 10732 &  7083.56 &     IGRINS &      240 &   110 \\
   HIP 11345 &  6894.90 &     CHIRON &      540 &   520 \\
   HIP 12332 &  6677.54 &       TS23 &     1107 &  1530 \\
   HIP 12332 &  6947.79 &     IGRINS &      224 &   170 \\
   HIP 12706 &  6584.86 &       TS23 &      362 &   990 \\
   HIP 12706 &  6946.84 &     IGRINS &      224 &   150 \\
   HIP 12719 &  6583.87 &       TS23 &      264 &  1720 \\
   HIP 12803 &  6676.53 &       TS23 &      452 &  1410 \\
   HIP 12803 &  6947.80 &     IGRINS &      224 &   140 \\
   HIP 12803 &  7240.95 &     IGRINS &      780 &   120 \\
   HIP 13165 &  6518.91 &        HRS &      275 &   850 \\
   HIP 13202 &  6910.83 &     CHIRON &      580 &   570 \\
   HIP 13209 &  6618.66 &       TS23 &      143 &  1880 \\
   HIP 13327 &  6670.54 &       TS23 &     1200 &  1800 \\
   HIP 13717 &  6917.83 &     CHIRON &      644 &   550 \\
   HIP 13717 &  6970.75 &     CHIRON &      703 &   550 \\
   HIP 13879 &  6945.79 &     IGRINS &      224 &   260 \\
   HIP 13879 &  6681.68 &       TS23 &      930 &   230 \\
   HIP 13879 &  6618.70 &       TS23 &      265 &  1490 \\
   HIP 14043 &  6523.92 &        HRS &      400 &   620 \\
   HIP 14143 &  6518.94 &        HRS &      462 &   900 \\
   HIP 14293 &  6919.89 &     CHIRON &      800 &   240 \\
   HIP 14576 &  6583.83 &       TS23 &       40 &  1030 \\
   HIP 14576 &  6582.92 &       TS23 &       36 &  2050 \\
   HIP 14764 &  6679.53 &       TS23 &      532 &  1370 \\
   HIP 14764 &  6946.80 &     IGRINS &      360 &   150 \\
   HIP 14862 &  6945.86 &     IGRINS &      224 &   180 \\
   HIP 14862 &  7084.60 &     IGRINS &     1200 &   120 \\
   HIP 15110 &  6946.83 &     IGRINS &      224 &   120 \\
   HIP 15338 &  6523.90 &        HRS &      570 &   690 \\
   HIP 15404 &  6517.92 &        HRS &      205 &   620 \\
   HIP 15444 &  6582.94 &       TS23 &      340 &  1790 \\
   HIP 16147 &  6299.59 &       TS23 &      735 &  1300 \\
   HIP 16210 &  6676.58 &       TS23 &      406 &  1440 \\
   HIP 16210 &  6945.84 &     IGRINS &      224 &   160 \\
   HIP 16244 &  6583.84 &       TS23 &      311 &  1750 \\
   HIP 16244 &  6945.83 &     IGRINS &      224 &   230 \\
   HIP 16285 &  6933.84 &     CHIRON &      767 &   570 \\
   HIP 16322 &  6944.72 &     CHIRON &      680 &   410 \\
   HIP 16322 &  6946.82 &     IGRINS &      224 &   170 \\
   HIP 16340 &  6678.54 &       TS23 &      570 &  1410 \\
   HIP 16599 &  6945.85 &     IGRINS &      224 &   170 \\
   HIP 16611 &  6946.85 &     IGRINS &      224 &   210 \\
   HIP 16611 &  6522.88 &     CHIRON &      255 &   580 \\
   HIP 17457 &  6531.82 &     CHIRON &     1005 &   520 \\
   HIP 17527 &  6947.82 &     IGRINS &      224 &   120 \\
   HIP 17527 &  6517.93 &        HRS &      429 &   670 \\
   HIP 17527 &  6963.91 &       TS23 &      818 &   470 \\
   HIP 17563 &  6538.82 &     CHIRON &     1088 &   400 \\
   HIP 18141 &  6678.59 &       TS23 &      605 &  1430 \\
   HIP 18396 &  6516.94 &        HRS &      259 &   510 \\
   HIP 18724 &  6299.62 &       TS23 &      226 &  1220 \\
   HIP 18788 &  6535.81 &     CHIRON &     1039 &   180 \\
   HIP 18805 &  6585.78 &       TS23 &      631 &  1760 \\
   HIP 19799 &  6668.63 &       TS23 &     1200 &  1900 \\
   HIP 19949 &  6945.90 &     IGRINS &      224 &   220 \\
   HIP 19949 &  7084.62 &     IGRINS &      960 &    30 \\
   HIP 19968 &  6676.62 &       TS23 &      524 &  1390 \\
   HIP 20264 &  6718.53 &     CHIRON &      860 &   410 \\
   HIP 20380 &  6945.89 &     IGRINS &      224 &   180 \\
   HIP 20380 &  7084.67 &     IGRINS &     1200 &    30 \\
   HIP 20430 &  6520.95 &        HRS &      360 &   710 \\
   HIP 20507 &  6719.50 &     CHIRON &      715 &   470 \\
   HIP 20579 &  6517.96 &        HRS &      490 &   830 \\
   HIP 20789 &  6946.86 &     IGRINS &      224 &   150 \\
   HIP 20789 &  6586.83 &       TS23 &      524 &  1610 \\
   HIP 21589 &  6962.94 &       TS23 &      220 &  1020 \\
   HIP 21589 &  7065.63 &     CHIRON &      754 &   290 \\
   HIP 21589 &  6948.96 &     IGRINS &      112 &   140 \\
   HIP 21683 &  6948.98 &     IGRINS &      230 &   110 \\
   HIP 21683 &  6962.97 &       TS23 &      335 &  1040 \\
   HIP 21819 &  6679.64 &       TS23 &      466 &  1480 \\
   HIP 21928 &  6677.63 &       TS23 &     1124 &  1470 \\
   HIP 22028 &  6947.90 &     IGRINS &      240 &   140 \\
   HIP 22509 &  6667.62 &       TS23 &     1200 &  1100 \\
   HIP 22509 &  7126.51 &     CHIRON &     2612 &   440 \\
   HIP 22833 &  6679.60 &       TS23 &      521 &  1520 \\
   HIP 22833 &  6946.87 &     IGRINS &      224 &   140 \\
   HIP 22840 &  6678.64 &       TS23 &      528 &  1380 \\
   HIP 22913 &  6585.74 &     CHIRON &     1200 &   230 \\
   HIP 22958 &  6551.79 &     CHIRON &     1200 &   340 \\
   HIP 22958 &  6943.77 &     CHIRON &      650 &   320 \\
   HIP 23362 &  6548.87 &     CHIRON &      725 &   520 \\
   HIP 23362 &  6946.93 &     IGRINS &      240 &   180 \\
   HIP 23916 &  6555.82 &     CHIRON &     1200 &    40 \\
   HIP 23916 &  6946.95 &     IGRINS &      280 &   140 \\
   HIP 24244 &  6949.84 &     CHIRON &      370 &   560 \\
   HIP 24244 &  6554.91 &     CHIRON &      488 &   330 \\
   HIP 24327 &  6529.88 &     CHIRON &      277 &   560 \\
   HIP 24505 &  7036.69 &     CHIRON &     2045 &   480 \\
   HIP 24902 &  6946.88 &     IGRINS &      224 &   180 \\
   HIP 25048 &  6299.69 &       TS23 &      900 &   420 \\
 ADS 3962 AB &  6946.92 &     IGRINS &      280 &   170 \\
   HIP 25143 &  7084.65 &     IGRINS &      960 &    80 \\
   HIP 25143 &  6946.89 &     IGRINS &      224 &   180 \\
   HIP 25280 &  6950.75 &     CHIRON &      597 &   540 \\
   HIP 25280 &  6947.91 &     IGRINS &      240 &   130 \\
   HIP 25281 &  6299.71 &       TS23 &      583 &  1120 \\
   HIP 25555 &  6670.63 &       TS23 &      600 &  1950 \\
   HIP 25608 &  6718.62 &     CHIRON &     1030 &   360 \\
   HIP 25695 &  6669.65 &       TS23 &     1200 &  1860 \\
   HIP 25790 &  6946.90 &     IGRINS &      224 &   160 \\
   HIP 25813 &  7083.58 &     IGRINS &     1020 &   200 \\
   HIP 25813 &  6544.88 &     CHIRON &      384 &   260 \\
   HIP 26063 &  6582.80 &     CHIRON &     1139 &   250 \\
   HIP 26093 &  6946.91 &     IGRINS &      280 &   150 \\
   HIP 26093 &  6676.66 &       TS23 &      525 &  1380 \\
   HIP 26093 &  7084.72 &     IGRINS &     1800 &    90 \\
   HIP 26126 &  6946.96 &     IGRINS &      240 &   170 \\
   HIP 26126 &  6721.50 &     CHIRON &      507 &   450 \\
   HIP 26126 &  6948.81 &     CHIRON &      550 &   360 \\
     HR 1880 &  6594.74 &     CHIRON &      472 &   540 \\
   HIP 26563 &  6677.68 &       TS23 &      502 &  1550 \\
   HIP 27100 &  6692.59 &     CHIRON &      294 &   680 \\
   HIP 27100 &  6727.54 &     CHIRON &      177 &   560 \\
   HIP 27321 &  6696.62 &     CHIRON &      178 &   640 \\
   HIP 27713 &  6946.97 &     IGRINS &      240 &   150 \\
   HIP 28691 &  6298.78 &       TS23 &      646 &  1300 \\
   HIP 28756 &  6571.74 &     CHIRON &     1200 &   320 \\
   HIP 28910 &  6693.61 &     CHIRON &      343 &   560 \\
   HIP 29150 &  6697.63 &     CHIRON &      950 &   380 \\
   HIP 29151 &  6946.99 &     IGRINS &      264 &   170 \\
   HIP 29151 &  6953.82 &     CHIRON &      617 &   540 \\
   HIP 29735 &  6559.80 &     CHIRON &      803 &   380 \\
   HIP 29735 &  6946.98 &     IGRINS &      240 &   170 \\
   HIP 29997 &  6679.70 &       TS23 &      415 &  1460 \\
   HIP 30069 &  6573.75 &     CHIRON &     1200 &   250 \\
   HIP 30073 &  6581.82 &     CHIRON &     1011 &   280 \\
   HIP 30666 &  6947.00 &     IGRINS &      280 &   160 \\
   HIP 30788 &  6574.75 &     CHIRON &      488 &   340 \\
   HIP 31278 &  7084.69 &     IGRINS &      960 &    90 \\
   HIP 31278 &  6668.71 &       TS23 &      797 &  1920 \\
   HIP 31278 &  6947.01 &     IGRINS &      280 &   170 \\
   HIP 31362 &  6598.72 &     CHIRON &     1200 &   250 \\
   HIP 31434 &  6676.71 &       TS23 &      611 &  1480 \\
   HIP 32474 &  6592.74 &     CHIRON &     1200 &   340 \\
   HIP 32607 &  7037.75 &     CHIRON &       56 &   210 \\
   HIP 32607 &  7039.62 &     CHIRON &       49 &   210 \\
   HIP 32607 &  6724.56 &     CHIRON &       67 &   560 \\
   HIP 33372 &  6948.00 &     IGRINS &      320 &   110 \\
   HIP 33372 &  6678.72 &       TS23 &      720 &  1380 \\
   HIP 33575 &  6599.72 &     CHIRON &     1200 &   390 \\
   HIP 34769 &  6677.75 &       TS23 &      279 &  1520 \\
   HIP 35180 &  6696.76 &     CHIRON &      925 &   290 \\
   HIP 35341 &  6680.75 &       TS23 &      833 &  1070 \\
   HIP 36393 &  6676.76 &       TS23 &      616 &  1550 \\
   HIP 36393 &  6670.76 &       TS23 &      494 &  1400 \\
   HIP 36760 &  7084.74 &     IGRINS &      720 &   150 \\
   HIP 36760 &  7037.76 &     CHIRON &     4284 &   370 \\
   HIP 36760 &  7039.68 &     CHIRON &     3127 &   480 \\
   HIP 36760 &  6668.79 &       TS23 &      836 &   700 \\
   HIP 36812 &  6947.99 &     IGRINS &      280 &   110 \\
   HIP 36917 &  6583.81 &     CHIRON &      571 &   420 \\
   HIP 36917 &  6971.83 &     CHIRON &      435 &   690 \\
     HR 2948 &  6584.81 &     CHIRON &      576 &   370 \\
   HIP 37297 &  6698.74 &     CHIRON &      370 &   480 \\
   HIP 37297 &  6574.82 &     CHIRON &      686 &   340 \\
   HIP 37297 &  6729.53 &     CHIRON &      231 &   460 \\
   HIP 37322 &  6975.79 &     CHIRON &      817 &   540 \\
   HIP 37450 &  6600.74 &     CHIRON &     1171 &   450 \\
   HIP 38538 &  6669.78 &       TS23 &      474 &  1990 \\
   HIP 38593 &  7039.70 &     CHIRON &      331 &   170 \\
   HIP 38593 &  6599.82 &     CHIRON &     1200 &   410 \\
   HIP 38593 &  6592.87 &     CHIRON &     1200 &   300 \\
   HIP 38846 &  6601.76 &     CHIRON &     1098 &   400 \\
   HIP 39095 &  6765.53 &     CHIRON &      430 &   450 \\
   HIP 39095 &  6757.56 &     CHIRON &      238 &   490 \\
   HIP 39095 &  6711.62 &     CHIRON &      288 &   500 \\
   HIP 39236 &  6679.73 &       TS23 &      730 &  1420 \\
   HIP 39567 &  6668.80 &       TS23 &      613 &  1940 \\
   HIP 39847 &  7084.68 &     IGRINS &      480 &   160 \\
   HIP 39847 &  6676.80 &       TS23 &      402 &  1500 \\
   HIP 39906 &  6580.86 &     CHIRON &      457 &   270 \\
   HIP 40429 &  6691.72 &     CHIRON &      710 &   560 \\
   HIP 40706 &  6327.64 &     CHIRON &      277 &   630 \\
   HIP 40881 &  6947.97 &     IGRINS &      240 &   100 \\
   HIP 41039 &  7039.73 &     CHIRON &     1728 &   530 \\
   HIP 41039 &  6330.77 &     CHIRON &      404 &   400 \\
   HIP 41307 &  6667.82 &       TS23 &      700 &  1220 \\
   HIP 42090 &  7124.62 &     IGRINS &      630 &   120 \\
   HIP 42129 &  6603.80 &     CHIRON &     1001 &   370 \\
   HIP 42313 &  6681.81 &       TS23 &      489 &  1020 \\
   HIP 42334 &  6712.64 &     CHIRON &      381 &   480 \\
   HIP 43142 &  6670.79 &       TS23 &      647 &  1980 \\
   HIP 44127 &  6677.88 &       TS23 &      239 &  1600 \\
   HIP 44307 &  6947.98 &     IGRINS &      280 &    80 \\
   HIP 45336 &  6328.70 &     CHIRON &      171 &   290 \\
   HIP 45336 &  7122.66 &     IGRINS &      270 &   130 \\
   HIP 45336 &  7084.75 &     IGRINS &      240 &   140 \\
   HIP 45344 &  6612.79 &     CHIRON &     1001 &   320 \\
   HIP 45688 &  6299.93 &       TS23 &      600 &   880 \\
   HIP 46225 &  6679.84 &       TS23 &      933 &  1300 \\
   HIP 46283 &  6732.56 &     CHIRON &      357 &   490 \\
   HIP 46283 &  6619.80 &     CHIRON &      872 &   430 \\
   HIP 46283 &  6712.70 &     CHIRON &      352 &   450 \\
   HIP 46897 &  7084.78 &     IGRINS &     1200 &    20 \\
   HIP 46897 &  7139.59 &     IGRINS &      720 &   120 \\
   HIP 47006 &  6677.03 &       TS23 &      373 &  1590 \\
   HIP 47175 &  6424.56 &     CHIRON &      264 &   660 \\
   HIP 50303 &  6677.78 &       TS23 &     1200 &  1350 \\
   HIP 50860 &  6678.81 &       TS23 &     1151 &  1470 \\
   HIP 51362 &  7043.79 &     CHIRON &     1298 &   340 \\
   HIP 51362 &  6625.81 &     CHIRON &      947 &   200 \\
   HIP 51362 &  6673.75 &     CHIRON &      606 &   510 \\
   HIP 51362 &  7084.76 &     IGRINS &      720 &   100 \\
   HIP 51685 &  6677.90 &       TS23 &     1046 &  1460 \\
   HIP 52422 &  6676.87 &       TS23 &      522 &  1540 \\
   HIP 52457 &  6676.83 &       TS23 &      647 &  1500 \\
   HIP 52638 &  6669.83 &       TS23 &      711 &  1900 \\
   HIP 52678 &  6748.56 &     CHIRON &      375 &   500 \\
   HIP 52678 &  6690.70 &     CHIRON &      618 &   610 \\
   HIP 52736 &  6748.52 &     CHIRON &      530 &   540 \\
   HIP 52736 &  6713.70 &     CHIRON &      263 &   490 \\
   HIP 52911 &  6670.86 &       TS23 &      758 &  1950 \\
     HR 4259 &  6335.70 &     CHIRON &      304 &   330 \\
   HIP 54849 &  6670.94 &       TS23 &      797 &  2080 \\
   HIP 55266 &  6300.00 &       TS23 &      900 &   200 \\
   HIP 55434 &  6332.77 &     CHIRON &      199 &   410 \\
   HIP 56034 &  6676.91 &       TS23 &      502 &  1460 \\
   HIP 56633 &  6669.91 &       TS23 &      447 &  1990 \\
   HIP 56633 &  6335.75 &     CHIRON &      358 &   470 \\
   HIP 57328 &  6338.68 &     CHIRON &      419 &   490 \\
   HIP 57328 &  6370.62 &     CHIRON &      419 &   450 \\
   HIP 58590 &  6338.72 &     CHIRON &      352 &   520 \\
   HIP 59394 &  6754.57 &     CHIRON &      815 &   500 \\
   HIP 59394 &  6690.78 &     CHIRON &      940 &   370 \\
   HIP 59449 &  6711.88 &     CHIRON &      186 &   440 \\
   HIP 59449 &  6641.76 &     CHIRON &      305 &   150 \\
   HIP 59819 &  6669.96 &       TS23 &      465 &  1940 \\
   HIP 60009 &  6370.66 &     CHIRON &      200 &   520 \\
   HIP 60009 &  7009.78 &     CHIRON &      206 &   510 \\
   HIP 60030 &  7084.81 &     IGRINS &     2400 &    80 \\
   HIP 60595 &  6676.94 &       TS23 &      746 &  1540 \\
   HIP 60710 &  7062.77 &     CHIRON &     1515 &   470 \\
   HIP 60957 &  6678.92 &       TS23 &      904 &  1500 \\
   HIP 61558 &  7139.67 &     IGRINS &     1440 &   180 \\
   HIP 61558 &  7084.84 &     IGRINS &     1200 &    70 \\
   HIP 61622 &  6378.69 &     CHIRON &      167 &   620 \\
   HIP 62541 &  7124.80 &     IGRINS &      480 &   120 \\
   HIP 62541 &  7084.86 &     IGRINS &     1500 &    80 \\
   HIP 62576 &  7122.83 &     IGRINS &      546 &   130 \\
   HIP 63724 &  6338.76 &     CHIRON &      412 &   560 \\
   HIP 63724 &  6371.72 &     CHIRON &      412 &   630 \\
   HIP 63945 &  6371.68 &     CHIRON &      362 &   560 \\
   HIP 65198 &  7044.85 &     CHIRON &     3352 &   430 \\
   HIP 65477 &  6670.01 &       TS23 &      207 &  2520 \\
   HIP 65728 &  6677.98 &       TS23 &     1200 &  1710 \\
   HIP 66249 &  6378.71 &     CHIRON &      110 &   520 \\
   HIP 66798 &  6377.82 &        HRS &      362 &   610 \\
   HIP 66821 &  6718.75 &     CHIRON &      610 &   540 \\
   HIP 67143 &  7139.77 &     IGRINS &      480 &   120 \\
   HIP 67194 &  6389.70 &        HRS &      368 &   740 \\
   HIP 67782 &  6394.92 &        HRS &      359 &   680 \\
   HIP 68092 &  7139.80 &     IGRINS &      840 &   200 \\
   HIP 68520 &  6403.59 &     CHIRON &      239 &   470 \\
   HIP 70327 &  7123.90 &     IGRINS &      420 &    90 \\
   HIP 70327 &  6681.98 &       TS23 &     1152 &   230 \\
   HIP 70327 &  6719.80 &     CHIRON &      368 &   460 \\
   HIP 70384 &  6403.73 &        HRS &      420 &   560 \\
   HIP 70400 &  6669.00 &       TS23 &      602 &  2000 \\
   HIP 70915 &  7136.63 &     CHIRON &     3281 &   480 \\
   HIP 70915 &  7090.71 &     CHIRON &     2985 &   460 \\
     HR 5476 &  6408.71 &        HRS &      305 &   660 \\
   HIP 71865 &  6383.71 &     CHIRON &      190 &   460 \\
   HIP 71865 &  7060.80 &     CHIRON &     1205 &   500 \\
   HIP 71974 &  7078.80 &     CHIRON &     3786 &   470 \\
   HIP 72104 &  6720.86 &     CHIRON &      492 &   490 \\
   HIP 72154 &  6403.75 &        HRS &      300 &   470 \\
   HIP 72250 &  7070.72 &     CHIRON &     4025 &   490 \\
   HIP 72378 &  7139.82 &     IGRINS &      720 &   140 \\
   HIP 72552 &  7124.81 &     IGRINS &      600 &   110 \\
   HIP 73049 &  6715.73 &     CHIRON &      749 &   530 \\
     HR 5605 &  6383.74 &     CHIRON &      330 &   530 \\
   HIP 74117 &  6383.77 &     CHIRON &      204 &   370 \\
   HIP 74689 &  6872.52 &     CHIRON &      693 &   550 \\
   HIP 75178 &  6678.99 &       TS23 &      815 &  1620 \\
   HIP 75304 &  6427.78 &     CHIRON &      309 &   340 \\
   HIP 76267 &  6380.88 &     CHIRON &       36 &   290 \\
   HIP 76600 &  6380.91 &     CHIRON &      137 &   500 \\
   HIP 76852 &  6426.65 &     CHIRON &      306 &   370 \\
   HIP 77233 &  6426.69 &     CHIRON &      140 &   320 \\
   HIP 77336 &  6396.79 &        HRS &      264 &   610 \\
   HIP 77336 &  7122.91 &     IGRINS &      480 &   120 \\
   HIP 77516 &  6380.89 &     CHIRON &      126 &   450 \\
   HIP 77635 &  6725.72 &     CHIRON &      288 &   420 \\
   HIP 77858 &  6733.75 &     CHIRON &      628 &   490 \\
   HIP 78105 &  6870.54 &     CHIRON &      655 &   590 \\
   HIP 78106 &  6736.70 &     CHIRON &      584 &   500 \\
   HIP 78265 &  6766.77 &     CHIRON &       22 &   250 \\
   HIP 78554 &  6427.74 &     CHIRON &      408 &   270 \\
   HIP 78820 &  6380.93 &     CHIRON &       53 &   490 \\
   HIP 78820 &  7139.83 &     IGRINS &      120 &   190 \\
   HIP 78820 &  7060.85 &     CHIRON &      263 &   360 \\
   HIP 78820 &  7156.63 &     CHIRON &      164 &   390 \\
   HIP 78821 &  7139.84 &     IGRINS &      210 &    50 \\
     HR 6025 &  6847.83 &     IGRINS &      600 &   100 \\
   HIP 78918 &  6718.86 &     CHIRON &      300 &   610 \\
   HIP 78933 &  6737.76 &     CHIRON &       77 &   250 \\
   HIP 79005 &  7090.77 &     CHIRON &     4646 &   430 \\
   HIP 79007 &  6874.48 &     CHIRON &      659 &   550 \\
   HIP 79199 &  6734.71 &     CHIRON &      451 &   470 \\
   HIP 79387 &  6746.75 &     CHIRON &      573 &   500 \\
   HIP 79404 &  6415.77 &     CHIRON &      324 &   470 \\
   HIP 79404 &  7076.81 &     CHIRON &     1307 &   430 \\
   HIP 79404 &  7090.88 &     CHIRON &     1135 &   450 \\
   HIP 79653 &  6740.79 &     CHIRON &      469 &   490 \\
   HIP 80460 &  7122.93 &     IGRINS &      720 &   110 \\
   HIP 80815 &  6378.86 &     CHIRON &      393 &   530 \\
   HIP 80883 &  6378.84 &     CHIRON &      174 &   550 \\
   HIP 80991 &  6389.86 &        HRS &      404 &   720 \\
   HIP 81126 &  7122.95 &     IGRINS &      720 &   150 \\
   HIP 81641 &  6404.83 &        HRS &      320 &   540 \\
   HIP 82350 &  6391.84 &        HRS &      404 &   660 \\
   HIP 82514 &  6726.81 &     CHIRON &       34 &   250 \\
   HIP 82673 &  6871.58 &     CHIRON &      250 &   310 \\
   HIP 82673 &  6410.86 &     CHIRON &      272 &   180 \\
   HIP 83635 &  6407.86 &        HRS &      316 &   480 \\
   HIP 84606 &  6945.54 &     IGRINS &      224 &   150 \\
   HIP 85290 &  6946.56 &     IGRINS &      224 &   120 \\
   HIP 85379 &  6398.85 &        HRS &      330 &   740 \\
   HIP 85385 &  6848.83 &     IGRINS &      480 &    80 \\
   HIP 85537 &  7096.88 &     CHIRON &     4227 &   570 \\
   HIP 85537 &  6792.73 &     CHIRON &      900 &   600 \\
   HIP 85537 &  6848.73 &     IGRINS &      240 &    60 \\
   HIP 85727 &  6771.85 &     CHIRON &      175 &   570 \\
   HIP 85922 &  6847.82 &     IGRINS &      480 &   120 \\
   HIP 85922 &  7139.95 &     IGRINS &      240 &   170 \\
   HIP 85922 &  6886.56 &     CHIRON &      791 &   590 \\
   HIP 85922 &  7103.85 &     CHIRON &     2543 &   370 \\
   HIP 85922 &  7104.82 &     CHIRON &     7386 &   360 \\
   HIP 85998 &  7122.96 &     IGRINS &      480 &    50 \\
   HIP 85998 &  7230.64 &     IGRINS &      720 &   170 \\
   HIP 85998 &  6407.84 &        HRS &      324 &   490 \\
   HIP 85998 &  7079.89 &     CHIRON &     1600 &   220 \\
   HIP 86019 &  6747.85 &     CHIRON &      228 &   510 \\
   HIP 86782 &  6404.85 &        HRS &      310 &   500 \\
   HIP 87108 &  6847.87 &     IGRINS &      120 &   120 \\
   HIP 87108 &  6445.79 &     CHIRON &      152 &   300 \\
   HIP 88116 &  6488.71 &     CHIRON &      373 &   590 \\
   HIP 88116 &  6876.54 &     CHIRON &      320 &   380 \\
   HIP 88290 &  6873.59 &     CHIRON &      365 &   540 \\
   HIP 88818 &  7122.98 &     IGRINS &      960 &    90 \\
   HIP 88818 &  7230.66 &     IGRINS &      600 &   110 \\
   HIP 88818 &  6397.85 &        HRS &      355 &   510 \\
   HIP 88817 &  7123.00 &     IGRINS &      960 &   100 \\
   HIP 88817 &  6405.84 &        HRS &      400 &   720 \\
   HIP 89156 &  6405.82 &        HRS &      470 &   610 \\
   HIP 89935 &  6848.85 &     IGRINS &      240 &   100 \\
   HIP 89935 &  6945.56 &     IGRINS &      224 &   250 \\
   HIP 90052 &  6405.88 &        HRS &      500 &   760 \\
   HIP 90762 &  6848.85 &     IGRINS &      480 &    80 \\
   HIP 90762 &  6945.57 &     IGRINS &      480 &   200 \\
   HIP 90887 &  6748.80 &     CHIRON &      710 &   560 \\
   HIP 91118 &  6945.58 &     IGRINS &      360 &   190 \\
   HIP 91875 &  6745.81 &     CHIRON &      680 &   580 \\
   HIP 92027 &  6405.91 &        HRS &      360 &   570 \\
   HIP 92312 &  6408.88 &        HRS &      390 &   650 \\
   HIP 92728 &  6405.86 &        HRS &      280 &   560 \\
   HIP 92855 &  6880.57 &     CHIRON &       30 &   550 \\
   HIP 92855 &  6415.80 &     CHIRON &       32 &   400 \\
   HIP 92946 &  6475.79 &     CHIRON &      339 &   500 \\
   HIP 93225 &  6946.54 &     IGRINS &      224 &    80 \\
   HIP 93225 &  6516.69 &     CHIRON &     1200 &   540 \\
   HIP 93225 &  6962.56 &       TS23 &      562 &  1430 \\
   HIP 93393 &  6404.88 &        HRS &      320 &   540 \\
   HIP 93580 &  7116.84 &     CHIRON &     5594 &   520 \\
   HIP 93580 &  6848.84 &     IGRINS &      480 &   100 \\
   HIP 93713 &  6848.87 &     IGRINS &      600 &    90 \\
   HIP 93747 &  6587.54 &       TS23 &       85 &  1880 \\
   HIP 93805 &  6852.82 &     IGRINS &     2640 &    60 \\
   HIP 93805 &  7139.96 &     IGRINS &      120 &   210 \\
   HIP 93805 &  6775.78 &     CHIRON &      150 &   520 \\
   HIP 94620 &  6848.88 &     IGRINS &      480 &   100 \\
   HIP 94720 &  6517.78 &        HRS &      400 &   880 \\
   HIP 95241 &  6869.63 &     CHIRON &      325 &   120 \\
   HIP 95241 &  6508.80 &     CHIRON &      294 &   600 \\
   HIP 95241 &  6459.89 &     CHIRON &      183 &   360 \\
   HIP 95560 &  6848.89 &     IGRINS &      480 &   110 \\
   HIP 95619 &  7139.97 &     IGRINS &      480 &   120 \\
   HIP 95619 &  7118.79 &     CHIRON &     3451 &   320 \\
   HIP 95619 &  6523.61 &     CHIRON &      768 &   670 \\
   HIP 95853 &  6585.54 &       TS23 &      129 &  2010 \\
   HIP 96288 &  6585.56 &       TS23 &      489 &  1860 \\
   HIP 96840 &  6408.92 &        HRS &      552 &   940 \\
   HIP 97376 &  6848.90 &     IGRINS &     1680 &    70 \\
   HIP 97496 &  6584.54 &       TS23 &      342 &  1820 \\
   HIP 97870 &  6517.81 &        HRS &      185 &   830 \\
   HIP 97966 &  6925.51 &     CHIRON &      655 &   380 \\
   HIP 97966 &  6881.70 &     CHIRON &     1158 &   660 \\
   HIP 98055 &  6586.54 &       TS23 &      320 &  1820 \\
   HIP 98194 &  6583.54 &       TS23 &      495 &  1750 \\
   HIP 98325 &  6587.56 &       TS23 &      543 &  1750 \\
   HIP 99080 &  6583.63 &       TS23 &      358 &  1740 \\
   HIP 99742 &  6945.59 &     IGRINS &      224 &   210 \\
  HIP 100069 &  6409.91 &        HRS &      451 &   560 \\
  HIP 100221 &  7230.72 &     IGRINS &     1100 &    80 \\
  HIP 100221 &  6523.79 &        HRS &      874 &   830 \\
  HIP 100881 &  6511.69 &     CHIRON &     1011 &   570 \\
  HIP 100907 &  6946.57 &     IGRINS &      224 &   120 \\
  HIP 100907 &  7230.67 &     IGRINS &      600 &   120 \\
  HIP 101123 &  6945.70 &     IGRINS &      224 &   160 \\
  HIP 101589 &  6448.83 &     CHIRON &      352 &   420 \\
  HIP 101716 &  6582.60 &       TS23 &      549 &  1260 \\
  HIP 101867 &  6489.78 &     CHIRON &      408 &   440 \\
  HIP 101909 &  6945.60 &     IGRINS &      360 &   140 \\
  HIP 101909 &  7230.69 &     IGRINS &      720 &    90 \\
  HIP 102487 &  6847.80 &     IGRINS &      480 &    80 \\
  HIP 103298 &  6946.58 &     IGRINS &      224 &   100 \\
  HIP 103298 &  7230.70 &     IGRINS &      600 &    80 \\
  HIP 103298 &  7155.93 &     CHIRON &     1141 &   290 \\
  HIP 104019 &  6474.90 &     CHIRON &      423 &   550 \\
  HIP 104105 &  6945.69 &     IGRINS &      280 &   140 \\
  HIP 104139 &  6448.90 &     CHIRON &      204 &   490 \\
  HIP 104365 &  6847.81 &     IGRINS &      480 &   100 \\
  HIP 104365 &  7237.82 &     IGRINS &      720 &    50 \\
  HIP 105140 &  6485.90 &     CHIRON &      372 &   420 \\
  HIP 105282 &  6522.65 &        HRS &     1006 &   850 \\
  HIP 105891 &  6946.59 &     IGRINS &      224 &    90 \\
  HIP 105942 &  6523.87 &        HRS &      480 &   760 \\
  HIP 105966 &  6585.64 &       TS23 &      530 &  1800 \\
  HIP 105972 &  6507.82 &        HRS &      267 &   690 \\
  HIP 106711 &  6584.60 &       TS23 &      689 &   700 \\
  HIP 106711 &  6618.53 &       TS23 &      557 &  2010 \\
  HIP 106786 &  6945.64 &     IGRINS &      224 &   230 \\
  HIP 106786 &  6962.59 &       TS23 &      284 &  1620 \\
  HIP 106786 &  6794.81 &     CHIRON &      232 &   470 \\
  HIP 107517 &  6873.74 &     CHIRON &     1025 &   490 \\
  HIP 107608 &  6788.84 &     CHIRON &      620 &   470 \\
  HIP 108294 &  6790.81 &     CHIRON &      470 &   510 \\
  HIP 108339 &  6945.65 &     IGRINS &      224 &   160 \\
  HIP 109056 &  6945.66 &     IGRINS &      240 &   150 \\
  HIP 109056 &  6948.56 &     IGRINS &     1156 &   160 \\
  HIP 109056 &  6946.65 &     IGRINS &      784 &   230 \\
  HIP 109056 &  6947.57 &     IGRINS &     1746 &   170 \\
  HIP 109139 &  6524.67 &     CHIRON &      207 &   590 \\
  HIP 109139 &  6447.91 &     CHIRON &      248 &   470 \\
  HIP 109521 &  6586.58 &       TS23 &      511 &  1860 \\
  HIP 109521 &  6962.60 &       TS23 &      587 &  1500 \\
  HIP 109521 &  6945.68 &     IGRINS &      224 &   200 \\
  HIP 109831 &  6945.67 &     IGRINS &      224 &   160 \\
  HIP 109831 &  7237.77 &     IGRINS &     1440 &   100 \\
  HIP 109831 &  7288.73 &     IGRINS &      900 &    30 \\
  HIP 110838 &  6543.82 &     CHIRON &      506 &   550 \\
  HIP 110838 &  6880.76 &     CHIRON &      354 &   780 \\
  HIP 110935 &  6896.67 &     CHIRON &      696 &   630 \\
  HIP 111056 &  7288.80 &     IGRINS &      180 &    30 \\
  HIP 111056 &  6945.76 &     IGRINS &      224 &   120 \\
  HIP 111068 &  6512.74 &        HRS &      362 &   500 \\
  HIP 111169 &  6585.62 &       TS23 &      121 &  1900 \\
  HIP 111841 &  6583.68 &       TS23 &      267 &  1670 \\
  HIP 112029 &  6487.85 &     CHIRON &      109 &   460 \\
  HIP 112029 &  6892.74 &     CHIRON &       95 &   250 \\
  HIP 113788 &  6586.63 &       TS23 &      367 &  1830 \\
  HIP 114520 &  6587.64 &       TS23 &      541 &  1880 \\
  HIP 114822 &  6886.68 &     CHIRON &      715 &   550 \\
  HIP 114822 &  6884.68 &     CHIRON &      530 &   530 \\
  HIP 115115 &  7163.90 &     CHIRON &      600 &   190 \\
  HIP 115115 &  7230.95 &     IGRINS &      500 &    40 \\
  HIP 115115 &  6797.87 &     CHIRON &      610 &   330 \\
  HIP 116247 &  6463.89 &     CHIRON &      368 &   560 \\
  HIP 116247 &  7237.94 &     IGRINS &      240 &   100 \\
  HIP 116582 &  6510.77 &        HRS &      534 &   670 \\
  HIP 116611 &  6946.61 &     IGRINS &      224 &   140 \\
  HIP 116631 &  6962.63 &       TS23 &      249 &  1980 \\
  HIP 116631 &  6946.78 &     IGRINS &      224 &   160 \\
  HIP 116631 &  6583.71 &       TS23 &      162 &  1740 \\
  HIP 116631 &  7288.80 &     IGRINS &      180 &    30 \\
  HIP 116805 &  6528.84 &       TS23 &      158 &  1680 \\
  HIP 116971 &  6527.85 &       TS23 &      407 &  1650 \\
  HIP 116971 &  6487.88 &     CHIRON &      298 &   540 \\
  HIP 116971 &  6555.56 &     CHIRON &      497 &   430 \\
  HIP 116971 &  6869.76 &     CHIRON &      380 &   110 \\
  HIP 116971 &  6936.73 &     CHIRON &      344 &   600 \\
  HIP 117089 &  7288.81 &     IGRINS &      180 &    20 \\
  HIP 117089 &  7237.95 &     IGRINS &      720 &    60 \\
  HIP 117089 &  6513.86 &     CHIRON &      853 &   600 \\
  HIP 117371 &  6586.67 &       TS23 &      349 &  1760 \\
  HIP 117452 &  6469.90 &     CHIRON &      327 &   540 \\
  HIP 118121 &  6795.86 &     CHIRON &      233 &   500 \\
  HIP 118243 &  6945.73 &     IGRINS &      224 &   170 \\
  HIP 118243 &  6583.74 &       TS23 &      351 &  1760
\enddata
\tablecomments{See Section \ref{subsec:specdata} for details of the signal-to-noise ratio calculation (the snr column)}
\label{tab:observations}
\end{deluxetable}

\clearpage

\begin{deluxetable*}{|l|llllllll|}
\tabletypesize{\small}
\tablewidth{0pt}
\tablecaption{NIRI Observation log}
\tablehead{
       \colhead{star} &  \colhead{K} &   \colhead{date} & \colhead{Exposure} & \colhead{$\rho$} & \colhead{$\theta$} & \colhead{$\Delta \mathrm{K_c}$} & \colhead{$M_2 (M_{\odot})$} & \colhead{$a$ (AU)} 
       }

\startdata
  HIP 22833 &  4.83 &  2015-03-29 &   50x0.3 & $0.352 +/- 0.003$ & $14.1 \pm 0.7$ & $3.04 \pm 0.06$ & $0.9^{+0.16}_{-0.17}$ & $25.1 \pm 0.24$ \\
  HIP 22958 &  5.79 &  2015-10-19 &   52x0.4 & $0.445 +/- 0.003$ & $65.0 \pm 0.3$ & $2.57 \pm 0.05$ & $1.5^{+0.35}_{-0.28}$ & $124.1 \pm 0.81$ \\
  HIP 26093 &  5.96 &  2015-03-23 &   33x0.6 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 26126 &  5.07 &  2015-03-29 &   40x0.4 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 29735 &  5.10 &  2015-03-22 &   83x0.2 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 31278 &  5.46 &  2015-03-28 &   29x0.6 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 38538 &  4.66 &  2015-10-19 &   75x0.2 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 39847 &  4.66 &  2015-03-27 &   50x0.3 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 75178 &  5.49 &  2015-03-24 &   29x0.6 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  HIP 88116 &  4.47 &  2015-03-22 &   50x0.3 & $3.671 +/- 0.001$ & $242.99 \pm 0.02$ & $3.9 \pm 0.4$ & $1.2^{+0.28}_{-0.24}$ & $467.7 \pm 0.18$ \\
  HIP 91118 &  5.67 &  2015-10-16 &   43x0.4 & $0.1667 +/- 0.0007$ & $41.7 \pm 0.2$ & $1.77 \pm 0.01$ & $1.5^{+0.39}_{-0.25}$ & $36.4 \pm 0.16$ \\
 HIP 100221 &  5.70 &  2015-10-18 &   26x0.7 & $0.250 +/- 0.002$ & $346.8 \pm 0.92$ & $2.28 \pm 0.04$ & $0.9^{+0.2}_{-0.23}$ & $28.5 \pm 0.24$ \\
 HIP 100907 &  5.41 &  2015-10-16 &   43x0.4 & $0.463 +/- 0.002$ & $335.4 \pm 0.30$ & $2.91 \pm 0.04$ & $0.7^{+0.14}_{-0.10}$ & $37.7 \pm 0.16$ \\
 HIP 101909 &  $6.47^{\dagger}$ &  2015-10-16 &   39x0.5 & \nodata & \nodata & \nodata & \nodata & \nodata \\
 HIP 109139 &  4.40 &  2015-04-26 &   50x0.3 & \nodata & \nodata & \nodata & \nodata & \nodata \\
 HIP 109521 &  4.96 &  2015-04-25 &   40x0.4 & \nodata & \nodata & \nodata & \nodata & \nodata \\
 HIP 115115 &  4.96 &  2015-10-18 &   40x0.4 & $1.497 +/- 0.002$ & $313.14 \pm 0.08$ & $2.54 \pm 0.04$ & $1.3^{+0.28}_{-0.19}$ & $120.0 \pm 0.17$
\enddata
\tablecomments{$\dagger$: There is no K-magnitude tabulated in the Simbad Database. The value quoted here is estimated from the spectral type of the star and its V-band magnitude.}
\label{tab:imaging_obs}
\end{deluxetable*}

\clearpage
\newpage

\LongTables
\begin{deluxetable*}{|l|rrlrrrll|}
\tabletypesize{\small}
\tablewidth{0pt}
\tablecaption{Companion Detections}
\tablehead{
     \colhead{} & \colhead{} & \colhead{} & \colhead{Used in} & \colhead{$T_\mathrm{eff}$} & \colhead{$v\sin{i}$} & \colhead{[Fe/H]} & \multicolumn{2}{c}{Mass ($M_{\odot}$)} \\ 
       \colhead{star} & \colhead{component} & \colhead{$N_\mathrm{obs}$} & \colhead{Analysis}  &   \colhead{(K)} &   \colhead{($\mathrm{km\ s}^{-1}$)} &  \colhead{(dex)} &   \colhead{Isochrone} &   \colhead{Spectral Type}
       }

\startdata
   HIP 2548 & Aa,Ab &     2 & yes &  $5732 \pm 112$ &      30 &    -0.5  &  $1.0^{+0.06}_{-0.04}$ &  $1.0^{+0.04}_{-0.03}$ \\
   HIP 5310 & AB &     1 & yes &  $3343 \pm 154$ &      30 &     0.0  &  $0.3^{+0.13}_{-0.08}$ &  $0.4^{+0.12}_{-0.12}$ \\
     HR 545 & Aa,Ab$^{\dagger}$ &     3 & yes &   $4312 \pm 87$ &       5 &    -0.5  &  $0.8^{+0.11}_{-0.10}$ &  $0.7^{+0.01}_{-0.02}$ \\
     HR 604 & AB &     1 & no &  $4736 \pm 150$ &       5 &     0.5  &    $0.9^{+0.17}_{-0.12}$ &  $0.8^{+0.02}_{-0.02}$ \\
  HIP 10732 & AB$^{\dagger}$ &     2 & yes &  $5578 \pm 109$ &       5 &     0.0  &  $1.0^{+0.29}_{-0.04}$ &  $1.0^{+0.02}_{-0.04}$ \\
  HIP 12332 & AB$^{\dagger}$ &     2 & yes &  $5551 \pm 107$ &      10 &     0.0  &    $1.0^{+0.03}_{-0.03}$ &  $1.0^{+0.01}_{-0.04}$ \\
  HIP 12706 & AB &     2 & yes &  $6051 \pm 107$ &      10 &     0.0  &  $1.1^{+0.03}_{-0.03}$ &  $1.1^{+0.04}_{-0.05}$ \\
  HIP 13165 & Aa,Ab &     1 & no &  $5773 \pm 163$ &       5 &    -0.5  &  $1.1^{+0.31}_{-0.08}$ &  $1.0^{+0.10}_{-0.03}$ \\
  HIP 13327 & AB$^{\dagger}$ &     1 & yes &  $5524 \pm 150$ &       5 &    -0.5  &  $1.2^{+0.20}_{-0.21}$ &  $1.0^{+0.03}_{-0.04}$ \\
  HIP 14576 & Aa,Ab &     2 & no &  $6980 \pm 103$ &      10 &     0.0  &  $1.4^{+0.03}_{-0.03}$ &  $1.5^{+0.05}_{-0.03}$ \\
  HIP 14764 & AB$^{\dagger}$ &     2 & yes &  $5623 \pm 150$ &      5 &     -0.5 &    $1.0^{+0.32}_{-0.07}$ &  $1.0^{+0.03}_{-0.04}$ \\
  HIP 16244 & AB &     2 & yes &  $5592 \pm 107$ &       5 &    -0.5  &  $1.2^{+0.20}_{-0.22}$ &  $1.0^{+0.03}_{-0.03}$ \\
  HIP 16340 & AB$^{\dagger}$ &     1 & yes &  $6214 \pm 150$ &       5 &    -0.5  &  $1.2^{+0.29}_{-0.07}$ &  $1.2^{+0.06}_{-0.05}$ \\
  HIP 18724 & AB &     1 & no &  $6831 \pm 154$ &      20 &     0.0  &  $1.7^{+0.13}_{-0.23}$ &  $1.4^{+0.04}_{-0.04}$ \\
  HIP 19949 & AB$^{\dagger}$ &     1 & yes &  $6492 \pm 155$ &      20 &    -0.5  &  $1.3^{+0.06}_{-0.05}$ &  $1.3^{+0.06}_{-0.06}$ \\
  HIP 20380 & AB$^{\dagger}$ &     2 & yes &  $6492 \pm 109$ &      20 &    -0.5  &  $1.3^{+0.04}_{-0.04}$ &  $1.3^{+0.03}_{-0.05}$ \\
  HIP 21589 & Aa,Ab &     1 & yes &  $4562 \pm 154$ &       5 &     0.0  &  $0.7^{+0.04}_{-0.03}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 22833 & AB &     2 & yes &  $4745 \pm 107$ &      10 &     0.0  &  $0.8^{+0.02}_{-0.02}$ &  $0.8^{+0.02}_{-0.01}$ \\
  HIP 22958 & Aa,Ab &     2 & yes &  $5762 \pm 158$ &      30 &    -0.5  &  $1.3^{+0.19}_{-0.23}$ &  $1.0^{+0.09}_{-0.03}$ \\
  HIP 23362 & AB$^{\dagger}$ &     1 & yes &  $6283 \pm 158$ &       5 &    -0.5  &  $1.2^{+0.06}_{-0.05}$ &  $1.2^{+0.07}_{-0.05}$ \\
  HIP 24902 & AB &     1 & yes &  $5679 \pm 154$ &      30 &     0.0  &  $1.0^{+0.05}_{-0.04}$ &  $1.0^{+0.04}_{-0.03}$ \\
  HIP 26126 & AB? &     3 & yes &   $5841 \pm 91$ &       5 &    -0.5  &  $1.0^{+0.03}_{-0.03}$ &  $1.0^{+0.05}_{-0.04}$ \\
  HIP 32607 & AB$^{\dagger}$ &     3 & yes &   $5068 \pm 91$ &       5 &    -0.5  &  $0.8^{+0.02}_{-0.02}$ &  $0.9^{+0.01}_{-0.01}$ \\
  HIP 38538 & AB      &     1 & no &   $6576 \pm 150$ &      30 &    -0.5  &  $1.3^{+0.05}_{-0.05}$ &  $1.3^{+0.09}_{-0.05}$ \\
  HIP 38593 & AB$^{\dagger}$ &     2 & yes &  $5398 \pm 112$ &       5 &    -0.5  &  $1.3^{+0.08}_{-0.13}$ &  $0.9^{+0.03}_{-0.03}$ \\
  HIP 38846 & AB$^{\dagger}$ &     1 & yes &  $6595 \pm 158$ &      10 &     0.0  &  $1.7^{+0.12}_{-0.21}$ &  $1.3^{+0.09}_{-0.05}$ \\
  HIP 39847 & Aa,Ab$^{\dagger}$ &     2 & no &  $5750 \pm 107$ &      10 &    -0.5  &  $1.0^{+0.04}_{-0.03}$ &  $1.0^{+0.05}_{-0.03}$ \\
  HIP 40881 & AB$^{\dagger}$ &     1 & yes &  $4562 \pm 154$ &      10 &    -0.5  &  $0.8^{+0.18}_{-0.05}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 42129 & AB &     1 & yes &  $5241 \pm 158$ &      10 &    -0.5  &  $1.2^{+0.12}_{-0.18}$ &  $0.9^{+0.04}_{-0.02}$ \\
  HIP 46283 & AB? &     3 & yes &   $9268 \pm 85$ &      10 &     0.5  &    $2.0^{+0.04}_{-0.02}$ &  $2.2^{+0.02}_{-0.02}$ \\
  HIP 52678 & AB$^{\dagger}$ &     2 & yes &  $5033 \pm 112$ &      10 &    -0.5  &  $1.0^{+0.15}_{-0.19}$ &  $0.8^{+0.01}_{-0.02}$ \\  
  HIP 55266 & AB &     1 & no &  $8483 \pm 139$ &      30 &    -0.5 &  $1.8^{+0.06}_{-0.04}$ &  $2.0^{+0.06}_{-0.05}$ \\
  HIP 63945 & AC &     1 & no &  $7038 \pm 162$ &      10 &    -0.5 &  $1.6^{+0.26}_{-0.16}$ &  $1.5^{+0.08}_{-0.06}$ \\
  HIP 72154 & AB$^{\dagger}$ &     1 & yes &  $4918 \pm 162$ &      30 &    -0.5  &  $0.9^{+0.22}_{-0.09}$ &  $0.8^{+0.03}_{-0.03}$ \\
  HIP 76267 & AB &     1 & yes &  $5449 \pm 158$ &       5 &    -0.5  &  $0.9^{+0.04}_{-0.04}$ &  $0.9^{+0.04}_{-0.03}$ \\
  HIP 77336 & AB$^{\dagger}$ &     2 & yes &  $6407 \pm 112$ &      20 &     0.0  &  $1.2^{+0.04}_{-0.04}$ &  $1.3^{+0.04}_{-0.05}$ \\
  HIP 77516 & AB &     1 & yes &  $6825 \pm 141$ &       5 &     0.5  &  $1.4^{+0.21}_{-0.06}$ &  $1.4^{+0.03}_{-0.04}$ \\
  HIP 78820 & AB &     1 & no &  $5658 \pm 158$ &       5 &    0 &  $1.3^{+0.11}_{-0.13}$ &  $1.0^{+0.04}_{-0.04}$ \\
  HIP 79199 & AB &     1 & no &  $4616 \pm 158$ &       5 &    -0.5  &  $0.9^{+0.14}_{-0.14}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 79404 & AB$^{\dagger}$ &     2 & yes &  $4773 \pm 112$ &      10 &    -0.5  &  $1.1^{+0.07}_{-0.08}$ &  $0.8^{+0.02}_{-0.01}$ \\
  HIP 80460 & AB &     1 & yes &  $6391 \pm 155$ &      10 &     0.0  &  $1.2^{+0.06}_{-0.05}$ &  $1.3^{+0.06}_{-0.06}$ \\
  HIP 82673 & AB$^{\dagger}$ &     2 & yes &  $4668 \pm 112$ &       5 &     0.5  &  $0.8^{+0.03}_{-0.03}$ &  $0.8^{+0.01}_{-0.02}$ \\
  HIP 84606 & AB &     1 & yes &  $5476 \pm 154$ &      10 &    -0.5 &  $0.9^{+0.05}_{-0.04}$ &  $0.9^{+0.04}_{-0.03}$ \\
  HIP 85998 & Aa,Ab &     3 & yes &  $5460 \pm 91$ &      1 &    -0.5  &  $1.0^{+0.28}_{-0.03}$ &  $0.9^{+0.04}_{-0.01}$ \\
  HIP 88116 & AB$^{\dagger}$ &     2 & yes &  $5293 \pm 112$ &       5 &     0.0  &  $0.9^{+0.03}_{-0.03}$ &  $0.9^{+0.02}_{-0.03}$ \\
  HIP 88817 & Ba,Bb$^{\dagger}$ &     1 & yes &  $3647 \pm 154$ &      20 &     0.0  &  $0.5^{+0.07}_{-0.10}$ &  $0.5^{+0.03}_{-0.05}$ \\
  HIP 88818 & Aa,Ab &     3 & yes &   $5296 \pm 91$ &      20 &    -0.5 &  $0.9^{+0.02}_{-0.02}$ &  $0.9^{+0.02}_{-0.03}$ \\
  HIP 91118 & AB &     1 & yes &  $6492 \pm 155$ &      10 &    -0.5  &  $1.3^{+0.26}_{-0.07}$ &  $1.3^{+0.06}_{-0.06}$ \\
  HIP 93805 & Aa,Ab$^{\dagger}$ &     2 & yes &  $4054 \pm 109$ &       5 &    -0.5  &  $0.6^{+0.03}_{-0.03}$ &  $0.6^{+0.04}_{-0.02}$ \\
  HIP 96840 & AB &     1 & yes &  $9942 \pm 152$ &      10 &     0.0  &    $2.2^{+0.06}_{-0.05}$ &  $2.3^{+0.03}_{-0.01}$ \\
 HIP 100221 & AB &     1 & yes &  $5273 \pm 154$ &      20 &    -0.5  &  $0.9^{+0.29}_{-0.07}$ &  $0.9^{+0.04}_{-0.03}$ \\
 HIP 100907 & AB$^{\dagger}$ &     2 & yes &  $5222 \pm 109$ &      10 &     0.0  &  $0.9^{+0.03}_{-0.03}$ &  $0.9^{+0.04}_{-0.02}$ \\
 HIP 103298 & Aa,Ab &     3 & yes &   $4615 \pm 90$ &       5 &    -0.5  &  $0.7^{+0.02}_{-0.02}$ &  $0.8^{+0.01}_{-0.02}$ \\
 HIP 109139 & AB &     2 & yes &  $5710 \pm 112$ &      20 &    -0.5  &  $1.0^{+0.04}_{-0.03}$ &  $1.0^{+0.03}_{-0.03}$ \\
 HIP 109521 & Aa,Ab$^{\dagger}$ &     3 & yes &   $4412 \pm 87$ &      10 &     0.0  &  $0.7^{+0.02}_{-0.02}$ &  $0.7^{+0.01}_{-0.01}$ \\
 HIP 115115 & AB &     3 & yes &   $4943 \pm 91$ &       5 &    -0.5  &  $0.8^{+0.02}_{-0.02}$ &  $0.8^{+0.02}_{-0.02}$

 \enddata

 \tablecomments{$\dagger$: Newly discovered companions.}
 \label{tab:companions}
\end{deluxetable*}

\clearpage
\newpage

\newpage
\clearpage
\bibliography{references}

\end{document}